---
title: "SLF Spread Through Time"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{vignette-060-spread-figure}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: inline
bibliography: references.bib
link-citations: yes
---

```{r, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.align='left'
)
```

Map of SLF Spread Through Time in the U.S.

***

We read in the packages to access data and generate the spread map. We also set the option to not download counties shapefiles every time (only the first).

# Setup

```{r setup, warning=FALSE, results='hide', message=FALSE}
library(slfrsk) #this package, has extract_enm()
library(tidyverse)  #data manipulation
library(here) #making directory pathways easier on different instances
library(sf) #simple features for easy shapefile manipulation
library(magrittr) #fast pipe assign
library(RColorBrewer) #palette
```

# Read in SLF reporting data

The data used here for the spread figure are based on an upcoming suite of packages (`lycodata` and `lycormap`) produced by the iEcoLab. The records in these packages contain data that are not yet available to the public, so all questions regarding the raw data should be directed to the creator of the `lycor*` suite, Seba De Bona (SDB, ([sebastiano.debona@temple.edu](mailto:sebastiano.debona@temple.edu))) or to Matthew R. Helmus ([mrhelmus@temple.edu](mailto:mrhelmus@temple.edu)). 

Here, we will read in a `.rda` file that was created using the example code in the Trade Relationship (`vignette-042-trade-presence`) vignette from the raw data.

```{r read in rda file}
#data call
data(by_county, package = "slfrsk")
```

# Plot spread figure

The data are read in above, and the year columns are also transformed into factors for easier plotting. 

```{r transform years to factors}
# transforming records and establishment
by_county %<>% 
  mutate(YearOfEstablishment = as.factor(YearOfEstablishment),
         FirstRecord = as.factor(FirstRecord))
```

Lastly, we plot the data after identifying a set of focal counties to outline (this bypasses layering issues with `sf` objects and plotting with `ggplot2`).

```{r plot spread map, fig.dim=c(8,10)}
#get the focal counties to outline
focal_counties <- by_county %>%
  filter(!is.na(FirstRecord)) %>%
  as.data.frame(.) %>%
  dplyr::select(GEOID) %>%
  unlist()

map_plot <- ggplot() +
  #geom_sf(data = by_county %>% filter(!GEOID %in% focal_counties), fill = "#FFFFFF", size = 0.5) +
    geom_polygon(data = map_data('state'), aes(x = long, y = lat, group = group), fill = "#FFFFFF", color = "black", lwd = 0.10) +
  geom_polygon(data = map_data('state', c("Connecticut","Delaware", "Maryland", "New Jersey", "New York", "Ohio", "Pennsylvania", "Virginia", "West Virginia")), aes(x = long, y = lat, group = group), fill = NA, color = "#e31a1c", lwd = 0.75) +
  geom_sf(data = by_county %>% filter(GEOID %in% focal_counties), aes(color = FirstRecord, fill = YearOfEstablishment), show.legend = T) +
  coord_sf(xlim = c(-124.7628, -66.94889), ylim = c(24.52042, 49.3833), expand = FALSE) +
  labs(x = "", y = "") +
  theme_bw() +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_rect(fill = "aliceblue")) +
  scale_color_brewer(palette = "YlOrRd", direction = 1, name = "Regulatory Incident", guide = guide_legend(override.aes = list(fill = "#808080"))) +
  scale_fill_brewer(palette = "YlOrRd", direction = 1, name = "Establishment", na.value = "#808080", breaks = c(2014,2015,2016,2017,2018,2019,2020)) +
  ggtitle(label = "")

map_plot
```

# Save plot output

```{r save spread plot for pub, eval=FALSE}
pdf(file.path(here::here(),"vignettes", paste0("spread_figure_v3.pdf")), width = (18/2.54), height = (14/2.54))
map_plot
invisible(dev.off())
```

# Extras

We also include a modification of SDB code here to show how to make a nice `leaflet` interactive map too... but we do not use this code.
```{r optional leaflet plotting, message=FALSE, eval=TRUE, fig.dim=c(8,10)}
library(leaflet)

colpal <- colorFactor(c(brewer.pal(7, "YlOrRd")),
                   by_county$YearOfEstablishment,
                   na.color = "transparent")
fillpal <- colorFactor(c(brewer.pal(7, "YlOrRd")),
                   by_county$FirstRecord,
                   na.color = "transparent")

leaflet(data = by_county, 
        options = leafletOptions(minZoom = 4, maxZoom = 18)) %>% 
  addProviderTiles(providers$CartoDB.Positron) %>% 
  addPolygons(label = ~NAME,
              color = ~colpal(FirstRecord),
              weight = 2,
              smoothFactor = 0.5,
              opacity = 0.5,
              fillOpacity = 0.5,
              fillColor = ~fillpal(YearOfEstablishment)) %>% 
  setView(lat = 40.80, lng = -100, zoom = 4) %>% 
  addLegend("bottomleft", pal = fillpal,
    title = "Invaded since",
    values = ~YearOfEstablishment,
    opacity = 1
  )

```
