---
title: "Wine Market Shock Severity Evaluation"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{vignette-041-market-plot}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: inline
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Build and Save Paninvasion Shock to Wine Market Plots

***

This vignette aims to reproduce the published figure(s) of the relationship between the predicted production of **grapes** from a severity model built from the alignment of the three risk potentials (**impact** potential regressed on **transport** and **establishment**) to the **wine market share** by country.

# Setup

We will first load the required packages to reproduce the published market risk figure.

```{r setup, warning=FALSE, results='hide', message=FALSE}
library(slfrsk)
library(tidyverse)
library(ggrepel)
library(DHARMa)
library(plotly)
```

# Load and summarize data

## Market size data

Next, we read in the `countries_market` dataset, which contains exports of wine by `country`. The data are in thousands of USD (1000's USD), and so we will need to convert to USD as we summarize the data. We summarize the `market` data by obtaining the **mean** of $log_{10}(value+1)$ transformed export values from 2012--2017.

We quickly visualize the distribution of these summarized `market_size` data via a histogram:

```{r FAO wine market, eval = TRUE, fig.height= 6, fig.width=7, warning=FALSE, message=FALSE}
data("countries_market")
#dim(countries_market)

# http://www.fao.org/faostat/en/#data/TM
countries_wine_market <- countries_market %>%
  filter(Year %in% (2012:2017)) %>%
  filter(Element == "Export Value") %>%
  group_by(`Area Code`,	Area, Year) %>%
  dplyr::select(`Area Code`,	Area, Year, Value) %>% 
  summarise(market_size = sum(Value)) %>%
  ungroup(Year) %>%

  mutate(market_size = log10((market_size*1000)+1)) %>% # units are in 1000s of USD
  #mutate(market_size = (market_size*1000)) %>%

  summarise(market_size = mean(market_size))  %>% 
  ungroup()

# remove all that are zero

countries_wine_market$market_size[countries_wine_market$market_size==0] <- NA
#countries_wine_market$market_size <- round(countries_wine_market$market_size)

#dim(countries_wine_market)
#countries_wine_market

qplot((countries_wine_market$market_size))
```

From here, we add an `ID` column that matches the other use of this label in the other package vignettes. Then, we isolate `ID` and `market_size` for severity analyses.

```{r add the ID column for joining to summary data, eval = TRUE}
#add the ID column for joining and plotting later
countries_wine_market <- countries_wine_market %>%
  mutate(ID = Area) %>%
  dplyr::select(ID, market_size) 
```

## Market summary data

To merge the new `market_size` dat with our existing `countries` data, we use `left_join()` to merge the `market` and `summary` data by `ID`. We then remove the countries from the dataset that already have **SLF**. Note: we use a modified version of `%in%` and `Negate()` to create `%notin%` for removing countries.

```{r join market to countries_summary_present_ensemble, eval = TRUE, message=FALSE}
#data("countries_summary_present")
data("countries_summary_present_ensemble")
countries_summary_present_ensemble <- countries_summary_present_ensemble %>%
    filter(!ID %in% c("Antarctica", "Monaco", "Norfolk Island", "Spratly Islands"))

#remove non-focal countries
rems <- c("Akrotiri and Dhekelia"
            ,"?land"
            ,"American Samoa"
            ,"Bouvet Island"
            ,"British Indian Ocean Territory"
            ,"Caspian Sea"
            ,"Christmas Island"
            ,"Clipperton Island"
            ,"Cocos Islands"
            ,"Falkland Islands"
            ,"Faroe Islands"
            ,"French Southern Territories"
            ,"Heard Island and McDonald Islands"
            ,"Mayotte"
            ,"Northern Mariana Islands"
            ,"Paracel Islands"
            ,"Pitcairn Islands"
            ,"Saint Pierre and Miquelon"
            ,"Tokelau"
            ,"United States Minor Outlying Islands"
            ,"Wallis and Futuna"
            ,"United States"
            ,"Philippines")

countries_summary_present_ensemble <- countries_summary_present_ensemble %>% filter(!ID %in% rems)


#dim(countries_summary_future)
`%notin%` <- Negate(`%in%`)

# filter(countries_summary_future, status == "native")

countries_sev_market <- countries_summary_present_ensemble %>% 
  left_join(countries_wine_market, by = "ID") %>%
  filter(ID %notin% c("United States", 
                      "China", 
                      "India", 
                      "Taiwan", 
                      "Japan", 
                      "South Korea", 
                      "Vietnam"))#%>% 

#########
# - Use this code to check to make sure that each country only has one row
# dim(countries_sev_market)
# setdiff(countries_sev_market$ID,countries_summary_future$ID)
# setdiff(countries_summary_future$ID,countries_sev_market$ID)
# table(countries_sev_market$ID)[table(countries_sev_market$ID)>1]
# table(countries_summary_future$ID)[table(countries_summary_future$ID)>1]

```

# Build and check model

From here, we now can build (or rebuild, technically, `vignette-021-quadrant-plots` does this first part as well) the severity model (`sev_mod`). This model is a linear regression model of log-transformed **impact** potential on the log-transformed **transport** and normal **establishment** potentials. We then evaluated the strength of the correlation between the empirical and predicted **impact** potential before conducting some diagnostics.

We then rescaled the predicted **impact** values to be bounded $[1,10]$ to produce initial `sev` values.

To create the initial *shock model* (`shock_mod`) and ultimately obtain a scaled measure of severity of **SLF** to shock the global wine market, we performed linear regression of the `market_size` data on `sev` for countries and conducted diagnostics as for `sev_mod`. We then evaluated the unscaled and scaled correlations between `market_size` and `sev` for countries, resulting in a severity of $8$ based on a scale of $[1,10]$.

```{r, fig.height= 7, fig.width=7, message=FALSE}
sev_mod <- lm(log10_avg_prod ~ log10_avg_infected_mass + grand_mean_max, 
               data = countries_sev_market)
cor.test(countries_sev_market$log10_avg_prod, predict(sev_mod), method = "spearman")

## diagnostics
# fit is good for the multivariate correlation because there is no dispersion, but do not use this model for prediction (which I do not know why you would)
simulationOutput_r <- simulateResiduals(fittedModel = sev_mod, plot = T) 
testDispersion(simulationOutput_r) 
summary(sev_mod)

countries_sev_market <- countries_sev_market %>%
  mutate(sev = scales::rescale(predict(sev_mod), to = c(1,10)))

shock_mod <- lm(market_size ~ sev, data = countries_sev_market)
summary(shock_mod)

## diagnostics
# severity estimate model fit is good and no major issues in heteroskedasticity. But I am not certain why you want this model anyways, a correlation is better because it is easily scaled between 1-10 
simulationOutput <- simulateResiduals(fittedModel = shock_mod, plot = T) 
testDispersion(simulationOutput)

cor.test(countries_sev_market$market_size, countries_sev_market$sev)
(shock_cor <- cor(countries_sev_market$market_size, 
                 countries_sev_market$sev,
                 use = "complete.obs") %>% 
  scales::rescale(to = c(1,10), from = c(-1,1)) %>%
  round())
```

# Plotting severity vs market scatter

With the predicted **severity**, we were able plot the scaled `sev` and `market_size`, resulting in the published severity of **SLF** to shock the global wine market plot.

## Static plot

```{r make scatter, eval = TRUE, fig.height= 12, fig.width=10, message=FALSE, warning=FALSE}
market_scatter <- countries_sev_market %>%
  
  ggplot(mapping = aes(x = (sev), 
                       y = (market_size), 
                       size = avg_wine, 
                       fill = wine)
         ) + 
  geom_abline(slope = 1, intercept = 0, lty = 2) +
  geom_abline(slope = 1, intercept = 5, lty = 2) +
  geom_abline(slope = 1, intercept = -5, lty = 2) +

  ggtitle(label = paste('Spotted Lanternfly Paninvasion Severity =',
                        shock_cor,'of 10 (P<0.001)')) +
  geom_point(pch = 21) + 
     scale_x_continuous(
    name =  substitute(expr = paste('Spotted Lanternfly (',italic("Lycorma delicatula"),') Invasion Severity')), 
    breaks = 1:10, 
    limits = c(1,10)
    ) +
  scale_y_continuous("Wine Export Market Size, log10 USD", 
                     breaks = seq(1,10,1), 
                     limits = c(1,10)
                     ) +
  scale_size_continuous(
        #values = c(0,),
       name = "Wine Production",# "Wine Impact Potential",
       # labels = c("0 t", "1.0M t", "2.0M t", "3.0M t", "4.0M t")
        labels = c("0 t", "1E6 t", "2E6 t", "3E6 t", "4E6 t")

      ) +
 # scale_fill_discrete(guide = FALSE) +
   scale_fill_manual( 
         values = c("no" = "#ffffff", "yes" = "#C77CFF"),
         name = "",# "Wine Impact Potential",
         labels = c("wine producer", "")
       ) +
      theme(
        panel.grid.major = element_line(colour = "#f0f0f0"),
        panel.grid.minor = element_blank(),
        #panel.grid.major = element_blank(),
        #panel.border = element_rect(colour = "black", fill=NA, size=1),
        axis.line = element_line(colour = "black"),
        #legend.position = c(0.7, 0.25),
        #legend.position = "bottom",
        panel.background = element_blank(),
        plot.background = element_blank(),
        #legend.background = element_rect(colour = 'black', 
        #                                 fill = 'white', 
        #                                 linetype='solid'),
        axis.text = element_text(size = rel(1)),
        axis.title.x = element_text(hjust = .4),
        axis.title.y = element_text(hjust = .35),
        legend.title = element_text(face = "plain"),
        legend.key.size = unit(0.2, "cm"),
        legend.key = element_blank(),
        #plot.margin = unit(c(hh, -5, hh, hh), units = "line"),
        #axis.title = element_text(size = rel(1.3))
      ) +
  geom_text_repel(#data = select(countries_sev_market,
                      aes(#x = x_to_plot, 
                          #y = y_to_plot, 
                          label = ID),
                      #label = countries_sev_market$ID,
                      min.segment.length = 0,
                      #direction = "x",
                      #nudge_y = nudge.y,
                      #nudge_x = nudge.x
      show.legend = FALSE) #+
  #guides(
  #      fill = TRUE
  #      ) #,
      #   color = guide_legend(
      #     order = 3,
      #     override.aes = list(alpha = 1)
      #   ),
      #   size = guide_legend(    # Adjust size to edit legend grape production
      #     order = 2,            # circles.
      #     override.aes = list(size = c(1,3,6), alpha = 1)
      #   )
      # ) 

market_scatter
```

## Interactive plot

We also provide an interactive version of the risk plot with `plotly`.

```{r, fig.height= 8, fig.width=8, message=FALSE, warning=FALSE}
#build ggplot
p1 <- countries_sev_market %>%
  mutate_if(.predicate = is.numeric, .funs = ~round(x = ., digits = 2)) %>%
    ggplot(mapping = aes(x = sev, 
                       y = market_size, 
                       size = avg_wine, 
                       fill = wine
                       )
         ) + 
  geom_abline(slope = 1, intercept = 0, lty = 2) +
  geom_abline(slope = 1, intercept = 5, lty = 2) +
  geom_abline(slope = 1, intercept = -5, lty = 2) +
  geom_point(pch = 21, aes(text = ID)) + 
  scale_fill_manual( 
         values = c("no" = "#ffffff", "yes" = "#C77CFF"),
         name = "",
         labels = c("wine producer", "")
       ) +
  lims(x = c(1, 10), y = c(1, 10)) +
  theme(
        panel.grid.major = element_line(colour = "#f0f0f0"),
        panel.grid.minor = element_blank(),
        axis.line = element_line(colour = "black"),
        panel.background = element_blank(),
        plot.background = element_blank(),
        axis.text = element_text(size = rel(1)),
        axis.title.x = element_text(hjust = .4),
        axis.title.y = element_text(hjust = .35),
        legend.title = element_text(face = "plain"),
        legend.key.size = unit(0.2, "cm"),
        legend.key = element_blank(),
      )

#execute plotly
ggplotly(p = p1, tooltip = c("ID", "sev", "market_size", "wine"), dynamicTicks = T) %>%
layout(xaxis = 
           list(title = list(text = paste('Spotted Lanternfly',sprintf("(<i>Lycorma delicatula</i>)"),'Invasion Severity'), font = list(size = 11)), showspikes = T),
         yaxis = 
           list(title = list(text = "Wine Export Market Size, log10 USD", font = list(size = 11)), showspikes = T),
         title = 
           list(text = paste('Spotted Lanternfly Paninvasion Severity =', shock_cor,'of 10 (P<0.001)'),
                font = list(size = 12),
                x = 0.1,
                y = 0.99),
         showlegend = F
         )
```

# Save plot output

Here, you can set to save the scatter plot to `/vignettes` as a `.pdf` file by setting the `if` from `FALSE` to `TRUE`, as in the other vignettes. We also have a version of the data saved for the Google Earth Engine App data cleaning vignette.

```{r, save pdf of image, message=TRUE, eval=TRUE}
if(FALSE){
  pdf(file.path(here::here(),
              "vignettes", 
              paste0("market_scatter_v4_1.pdf")),
    width = 8.5*.95, height = 11*.95)

market_scatter
invisible(dev.off())

countries_sev <- countries_sev_market %>% 
  dplyr::select(ID, market_size, sev)

save(countries_sev, file = file.path(here::here(), "data", "countries_sev.rda"))
}
```

