---
title: "Invasion Potential Alignment ALT"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{vignette-025-quadrant-plots-alt}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Build and Save Risk Correlation Quadrant Plots (Ensemble Extracts, Alternate Extraction Data)

1. Load required packages

```{r load required packages, warning=FALSE, message=FALSE}
library(slfrsk)
library(tidyverse)
```

2. Load the full `suitability extract ensemble` data and the `summary_future` data (`states` and `countries`). Join the two dataframes and resolve conflicts. Use the `summary_future` data to lead the merging with `left_join()`.

  - mean suitability
  - median suitability
  
```{r read in data}
#suitability data load
data("states_extracts_ensemble")
data("countries_extracts_ensemble")

#summary data load
data("states_summary_future_ensemble")
data("countries_summary_future_ensemble")

#join the data
states_summary_future_ensemble <- states_summary_future_ensemble %>%
  left_join(., states_extracts_ensemble, by = "geopol_unit") %>%
  #drop grandmeanmax col, it is redundant
  dplyr::select(-grand_mean_max)

countries_summary_future_ensemble <- countries_summary_future_ensemble %>%
  left_join(., countries_extracts_ensemble, by = "geopol_unit") %>%
  #drop grandmeanmax col, it is redundant
  dplyr::select(-grand_mean_max)

```
