---
title: "vignette-010-tidy-data"
output: rmarkdown::html_notebook
vignette: >
  %\VignetteIndexEntry{vignette-010-tidy-data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}

---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Wrangle and Tidy Package data

***

All of the following data sources will be read in as raw data from `slfrsk/data/` (originally cleaned a bit from raw data versions in `slfrsk/data-raw/`) and then cleaned up individually in this vignette (see `/data-raw/convert_data_rda.R` for script and commenting that includes earlier unit conversion and transformation of raw data from `.csv` to `.rda` files used here). Depending on the situation, the resultant "cleaned/tidied" data may be saved to `slfrsk/data/` as a new `.rda` file. In this vignette, the following data are treated:
  
  1. Establishment potential
  2. Transport potential
  3. Impact potential

***

# Setup

Here we load the required packages to prepare the data.

```{r setup, warning=FALSE, results='hide', message=FALSE}
library(slfrsk)
library(tidyverse)
library(here)
devtools::load_all()
```

# Establishment
## Maxent Ensemble Model

Previously, we built three separate species distribution models (SDMs) with Maxent. Upstream of this vignette, we obtained summary statistics for suitability in each geopolitical unit (country and U.S. state) in each SDM, which were combined into the `states_extracts` and `countries_extracts` datasets. 

Here, we explore the ensemble *maximum suitability* by obtaining the maximum suitability per geopolitical unit for each SDM and calculating the *mean* and *standard error* across the models. These summarized suitabilities constitute our predicted **Establishment potential**.

We also strip *Western Sahara* from the `countries` datasets here and for other datsets, as it is not informative for our analyses (zero trade for **Transport potential** below). 
  
```{r bring in maxent extraction summary data, message=FALSE}
data("states_extracts")

summary_states_extracts <- states_extracts %>%
  group_by(geopol_unit) %>%
  summarize(grand_mean_max = mean(obs_max), grand_se_max = sd(obs_max) / sqrt(n_distinct(model)))

data("countries_extracts")
  
summary_countries_extracts <- countries_extracts %>%
  group_by(geopol_unit) %>%
  summarize(grand_mean_max = mean(obs_max), grand_se_max = sd(obs_max) / sqrt(n_distinct(model)))

#rm W. Sahara
summary_countries_extracts <- summary_countries_extracts %>%
    filter(geopol_unit != "Western Sahara")

```

***

# Transport
## Import Trade Data

Now that **Establishment** data are nicely organized, we can move on to **Transport**. We estimate **Transport potential** based on the average annual mass of goods traded (in metric tons) with *Lycorma*-established U.S. states over 2012--2017 (we also track value of traded goods in USD). Our existing data can be partitioned into geopolitical units as above (states and countries), but can further be separated into two categories: 1. states with confirmed _Lycorma_ populations (`present`) and 2. states that have a high probability of hosting _Lycorma_ populations in the near future based on proximity to `present` states and records of isolated observations (`future`). 

Although we maintain several possible permutations of data to analyze, we focus on mass of goods traded for `future` scenarios, as these data are unlikely to be skewed by trade of high value goods that are less likely to lead to transportation of egg masses (e.g., electronics) and provide us with a more careful outlook for risk of slf paninvasion.

### U.S. States

We first treat the states data by adding two letter codes and *Lycorma*-establishment status. For `present`, we treat the following states as having established *Lycorma*: **DE, MD, NJ, PA, and VA**. For `future`, we add the following states as having established *Lycorma*: **NC, NY, OH, and WV**. From here, the data are tidied and gathered before calculating the mean and standard error (SE) of trade with established states for 2012--2017. To do so, we first removed all cases of intrastate trade for established states. Then, we calculated the total trade with established states for each year before obtaining the overall mean and SE.

Below, we demonstrate the code for both `present` and `future` trade data in value (USD) for U.S. `states`.

```{r load in states trade data and clean value, message=FALSE, warning=FALSE}
#load states value
data("states_trade_summary_slf")
data("states_trade_summary_slf_future")

#code snippet to attach infection status and add state abbreviations
#state id's for adding to extracts
stateid <- c("AL", "AK", "AZ", "AR", "CA", "CO", "CT", "DE", "FL", "GA", "HI", "ID", "IL", "IN", "IA", "KS", "KY", "LA", "ME", "MD", "MA", "MI", "MN", "MS", "MO", "MT", "NE", "NV", "NH", "NJ", "NM", "NY", "NC", "ND", "OH", "OK", "OR", "PA", "RI", "SC", "SD", "TN", "TX", "UT", "VT", "VA", "WA", "DC", "WV", "WI", "WY")
stateid <- tibble(stateid, states_trade_summary_slf$destination)
colnames(stateid) <- c("ID", "geopol_unit")

#PRESENT

#add in the infection status of each state while adding the abbreviations
 states_trade_summary_slf <-  left_join(states_trade_summary_slf, stateid, by = c("destination" = "geopol_unit")) %>%
    mutate(status = ifelse(ID %in% c("DE", "MD", "NJ","PA", "VA"), "established", "not established")) %>%
    select(destination, ID, status, everything())
  
  #tidying of data
  states_trade_summary_slf <- states_trade_summary_slf %>%
    gather(-destination:-status, key = "infected_state", value = "trade") %>%
    as.data.frame(.)
  
  #total infected states trade, average/SE
  summary_states_trade_value <- states_trade_summary_slf %>%
    group_by(destination) %>%
    #filter to be just trade with infected states present
    filter(str_detect(infected_state, "pa_") | str_detect(infected_state, "de_") | str_detect(infected_state, "va_") | str_detect(infected_state, "nj_") | str_detect(infected_state, "md_")) %>%
    #remove the all state trade from consideration
    filter(str_detect(infected_state, "201") & !str_detect(infected_state, "all")) %>%  
    #make a temporary col
    mutate(tempcol = str_split(infected_state, "_")) %>%  
    #group by row
    rowwise() %>% 
    #split out the info about states and years of trade
    mutate(source = unlist(tempcol)[1], year = unlist(tempcol)[2]) %>%
    #rm the temp col
    select(-tempcol) %>% 
    #group based now on the destination, year, and infection status
    #filter out the intrastate trade for infected states to avoid extra low averages by creating a hybrid string (DestinationState+abbreviated_infected state)
    group_by(destination, year, status) %>% 
    mutate(clear_intra = paste0(destination, source)) %>%
    filter(!clear_intra %in% c("Delawarede", "Marylandmd", "New Jerseynj", "Pennsylvaniapa", "Virginiava")) %>%
    #remove workaround infected intrastate column
    dplyr::select(-clear_intra) %>%
    #add up the trade with infected states for each destination state and year (THIS IS NOT AVG INFECTED TRADE YET)
    summarize(avg_infected_trade = sum(trade)) %>%  
    group_by(destination, status) %>%
    #obtain the standard error (SE) of infected trade across years (this repeats across all of the entries across years and thus becomes its own value)
    mutate(se = (sd(avg_infected_trade, na.rm = T) / sqrt(length(year)))) %>% 
    #summarize the mean and SE across years! (there is no need to actually take the mean of SE here, but it makes summarizing simplier)
    summarise(avg_infected_trade = mean(avg_infected_trade, na.rm = T), se = mean(se))  
  
#FUTURE
    states_trade_summary_slf_future <-  left_join(states_trade_summary_slf_future, stateid, by = c("destination" = "geopol_unit")) %>%
    mutate(status = ifelse(ID %in% c("DE", "MD", "NC", "NJ", "NY", "OH", "PA", "VA", "WV"), "established", "not established")) %>%
    dplyr::select(destination, ID, status, everything())
  
  #tidying of data
  states_trade_summary_slf_future <- states_trade_summary_slf_future %>%
    gather(-destination:-status, key = "infected_state", value = "trade") %>%
    as.data.frame(.)
  
  #total infected states trade, average/SE
  summary_states_trade_value_future <- states_trade_summary_slf_future %>%
    group_by(destination) %>%
    filter(str_detect(infected_state, "pa_") | str_detect(infected_state, "de_") | str_detect(infected_state, "va_") | str_detect(infected_state, "nj_") | str_detect(infected_state, "md_") | str_detect(infected_state, "nc_")  | str_detect(infected_state, "ny_")  | str_detect(infected_state, "oh_")  | str_detect(infected_state, "wv_")) %>%
    filter(str_detect(infected_state, "201") & !str_detect(infected_state, "all")) %>%
    mutate(tempcol = str_split(infected_state, "_")) %>%
    rowwise() %>%
    mutate(source = unlist(tempcol)[1], year = unlist(tempcol)[2]) %>%
    select(-tempcol) %>%
    group_by(destination, year, status) %>%
    #filter out the intrastate trade for infected states to avoid extra low averages
    mutate(clear_intra = paste0(destination, source)) %>%
    filter(!clear_intra %in% c("Delawarede", "Marylandmd", "New Jerseynj", "Pennsylvaniapa", "Virginiava", "North Carolinanc", "New Yorkny", "Ohiooh", "West Virginiawv")) %>%
    #remove workaround infected intrastate column
    dplyr::select(-clear_intra) %>%
    summarize(avg_infected_trade = sum(trade)) %>%
    group_by(destination, status) %>%
    mutate(se = (sd(avg_infected_trade, na.rm = T) / sqrt(length(year)))) %>%
    summarise(avg_infected_trade = mean(avg_infected_trade, na.rm = T), se = mean(se))
```

We repeat the same process for `states` but for the total mass of imported goods (`mass`). In doing so, we convert the existing units from kg to **metric tons**.
    
```{r clean states mass, message=FALSE, warning=FALSE}
#load states mass
data("states_trade_mass_summary_slf")
data("states_trade_mass_summary_slf_future")

#PRESENT
  #add in the infection status of each state while adding the abbreviations
   states_trade_mass_summary_slf <- left_join(states_trade_mass_summary_slf, stateid, by = c("destination" = "geopol_unit")) %>%
    mutate(status = ifelse(ID %in% c("DE", "MD", "NJ","PA", "VA"), "established", "not established")) %>%
    select(destination, ID, status, everything())

     #tidying of data and dividing by 1000 to convert from kg to metric tons
  states_trade_mass_summary_slf <- states_trade_mass_summary_slf %>%
    gather(-destination:-status, key = "infected_state", value = "mass") %>%
    mutate(mass = mass / 1000) %>%
    as.data.frame(.)
  
  #mass
  summary_states_trade_mass <- states_trade_mass_summary_slf %>%
    group_by(destination) %>%
    filter(str_detect(infected_state, "pa_") | str_detect(infected_state, "de_") | str_detect(infected_state, "va_") | str_detect(infected_state, "nj_") | str_detect(infected_state, "md_")) %>%
    filter(str_detect(infected_state, "201") & !str_detect(infected_state, "all")) %>%
    mutate(tempcol = str_split(infected_state, "_")) %>%
    rowwise() %>%
    mutate(source = unlist(tempcol)[1], year = unlist(tempcol)[2]) %>%
    select(-tempcol) %>%
    group_by(destination, year, status) %>%
    #filter out the intrastate trade for infected states to avoid extra low averages
    mutate(clear_intra = paste0(destination, source)) %>%
    filter(!clear_intra %in% c("Delawarede", "Marylandmd", "New Jerseynj", "Pennsylvaniapa", "Virginiava")) %>%
    #remove workaround infected intrastate column
    dplyr::select(-clear_intra) %>%
    summarize(avg_infected_mass = sum(mass)) %>%
    group_by(destination, status) %>%
    mutate(se = (sd(avg_infected_mass, na.rm = T) / sqrt(length(year)))) %>%
    summarise(avg_infected_mass = mean(avg_infected_mass, na.rm = T), se = mean(se))

#FUTURE
   #add col to identify infected states
  states_trade_mass_summary_slf_future <-  left_join(states_trade_mass_summary_slf_future, stateid, by = c("destination" = "geopol_unit")) %>%
    mutate(status = ifelse(ID %in% c("DE", "MD", "NC", "NJ", "NY", "OH", "PA", "VA", "WV"), "established", "not established")) %>%
    dplyr::select(destination, ID, status, everything())
  
  #tidying of data and dividing by 1000 to convert from kg to metric tons
  states_trade_mass_summary_slf_future <- states_trade_mass_summary_slf_future %>%
    gather(-destination:-status, key = "infected_state", value = "mass") %>%
    mutate(mass = mass / 1000) %>%
    as.data.frame(.)
  
  #mass
  summary_states_trade_mass_future <- states_trade_mass_summary_slf_future %>%
    group_by(destination) %>%
    filter(str_detect(infected_state, "pa_") | str_detect(infected_state, "de_") | str_detect(infected_state, "va_") | str_detect(infected_state, "nj_") | str_detect(infected_state, "md_") | str_detect(infected_state, "nc_")  | str_detect(infected_state, "ny_")  | str_detect(infected_state, "oh_")  | str_detect(infected_state, "wv_")) %>%
    filter(str_detect(infected_state, "201") & !str_detect(infected_state, "all")) %>%
    mutate(tempcol = str_split(infected_state, "_")) %>%
    rowwise() %>%
    mutate(source = unlist(tempcol)[1], year = unlist(tempcol)[2]) %>%
    select(-tempcol) %>%
    group_by(destination, year, status) %>%
    #filter out the intrastate trade for infected states to avoid extra low averages
    mutate(clear_intra = paste0(destination, source)) %>%
    filter(!clear_intra %in% c("Delawarede", "Marylandmd", "New Jerseynj", "Pennsylvaniapa", "Virginiava", "North Carolinanc", "New Yorkny", "Ohiooh", "West Virginiawv")) %>%
    #remove workaround infected intrastate column
    dplyr::select(-clear_intra) %>%
    summarize(avg_infected_mass = sum(mass)) %>%
    group_by(destination, status) %>%
    mutate(se = (sd(avg_infected_mass, na.rm = T) / sqrt(length(year)))) %>%
    summarise(avg_infected_mass = mean(avg_infected_mass, na.rm = T), se = mean(se))
```

> Here, you can see the summary for the first 10 states for the `future` trade data for mass of trade (metric tons).    
    
```{r show the data format for summarized trade so far for states, echo=FALSE}
head(summary_states_trade_mass_future, n = 10)
```

### Countries

From here, we turn to complete the same analyses for `countries` as states (value of trade in USD and then mass of trade in metric tons). Notably, we do not have to worry about the intrastate trade that we filtered out in the `states` datasets. As such, we do not bother with attaching a `$status` identifier to each country (this is done as needed for visualization later).

```{r load in countries trade data and clean value, message=FALSE,warning=FALSE}
#COUNTRIES

#VALUE
#load countries values
data("countries_trade_summary_slf")
data("countries_trade_summary_slf_future")

#PRESENT

#tidying of data and log transforming values
countries_trade_summary_slf <- countries_trade_summary_slf %>%
  gather(-destination, key = "infected_state", value = "trade") %>%
  as.data.frame(.)

  #value
  summary_countries_trade_value <- countries_trade_summary_slf %>%
    group_by(destination) %>%
    filter(str_detect(infected_state, "pa_") | str_detect(infected_state, "de_") | str_detect(infected_state, "va_") | str_detect(infected_state, "nj_") | str_detect(infected_state, "md_")) %>%
    filter(str_detect(infected_state, "201") & !str_detect(infected_state, "all")) %>%
    mutate(tempcol = str_split(infected_state, "_")) %>%
    rowwise() %>%
    mutate(source = unlist(tempcol)[1], year = unlist(tempcol)[2]) %>%
    select(-tempcol) %>%
    group_by(destination, year) %>%
    summarize(avg_infected_trade = sum(trade)) %>%
    group_by(destination) %>%
    mutate(se = (sd(avg_infected_trade, na.rm = T) / sqrt(length(year)))) %>%
    summarise(avg_infected_trade = mean(avg_infected_trade, na.rm = T), se = mean(se))

#FUTURE
  
#tidying of data and log transforming values
countries_trade_summary_slf_future <- countries_trade_summary_slf_future %>%
  gather(-destination, key = "infected_state", value = "trade") %>%
  as.data.frame(.)

  #value
  summary_countries_trade_value_future <- countries_trade_summary_slf_future %>%
    group_by(destination) %>%
    filter(str_detect(infected_state, "pa_") | str_detect(infected_state, "de_") | str_detect(infected_state, "va_") | str_detect(infected_state, "nj_") | str_detect(infected_state, "md_") | str_detect(infected_state, "nc_")  | str_detect(infected_state, "ny_")  | str_detect(infected_state, "oh_")  | str_detect(infected_state, "wv_")) %>%
    filter(str_detect(infected_state, "201") & !str_detect(infected_state, "all")) %>%
    mutate(tempcol = str_split(infected_state, "_")) %>%
    rowwise() %>%
    mutate(source = unlist(tempcol)[1], year = unlist(tempcol)[2]) %>%
    select(-tempcol) %>%
    group_by(destination, year) %>%
    summarize(avg_infected_trade = sum(trade)) %>%
    group_by(destination) %>%
    mutate(se = (sd(avg_infected_trade, na.rm = T) / sqrt(length(year)))) %>%
    summarise(avg_infected_trade = mean(avg_infected_trade, na.rm = T), se = mean(se))

#MASS
#load countries mass
data("countries_trade_mass_summary_slf")
data("countries_trade_mass_summary_slf_future")

#PRESENT

#tidying of data and converting from kg to metric tons 
countries_trade_mass_summary_slf <- countries_trade_mass_summary_slf %>%
  gather(-destination, key = "infected_state", value = "mass") %>%
  mutate(mass = mass / 1000) %>%
  as.data.frame(.)

  #mass
  summary_countries_trade_mass <- countries_trade_mass_summary_slf %>%
    group_by(destination) %>%
    filter(str_detect(infected_state, "pa_") | str_detect(infected_state, "de_") | str_detect(infected_state, "va_") | str_detect(infected_state, "nj_") | str_detect(infected_state, "md_")) %>%
    filter(str_detect(infected_state, "201") & !str_detect(infected_state, "all")) %>%
    mutate(tempcol = str_split(infected_state, "_")) %>%
    rowwise() %>%
    mutate(source = unlist(tempcol)[1], year = unlist(tempcol)[2]) %>%
    select(-tempcol) %>%
    group_by(destination, year) %>%
    summarize(avg_infected_mass = sum(mass)) %>%
    group_by(destination) %>%
    mutate(se = (sd(avg_infected_mass, na.rm = T) / sqrt(length(year)))) %>%
    summarise(avg_infected_mass = mean(avg_infected_mass, na.rm = T), se = mean(se))
  
#FUTURE
  
#tidying of data and converting from kg to metric tons 
countries_trade_mass_summary_slf_future <- countries_trade_mass_summary_slf_future %>%
  gather(-destination, key = "infected_state", value = "mass") %>%
  mutate(mass = mass / 1000) %>%
  as.data.frame(.)

  #mass
  summary_countries_trade_mass_future <- countries_trade_mass_summary_slf_future %>%
    group_by(destination) %>%
    filter(str_detect(infected_state, "pa_") | str_detect(infected_state, "de_") | str_detect(infected_state, "va_") | str_detect(infected_state, "nj_") | str_detect(infected_state, "md_") | str_detect(infected_state, "nc_")  | str_detect(infected_state, "ny_")  | str_detect(infected_state, "oh_")  | str_detect(infected_state, "wv_")) %>%
    filter(str_detect(infected_state, "201") & !str_detect(infected_state, "all")) %>%
    mutate(tempcol = str_split(infected_state, "_")) %>%
    rowwise() %>%
    mutate(source = unlist(tempcol)[1], year = unlist(tempcol)[2]) %>%
    select(-tempcol) %>%
    group_by(destination, year) %>%
    summarize(avg_infected_mass = sum(mass)) %>%
    group_by(destination) %>%
    mutate(se = (sd(avg_infected_mass, na.rm = T) / sqrt(length(year)))) %>%
    summarise(avg_infected_mass = mean(avg_infected_mass, na.rm = T), se = mean(se))
```

> Here, you can see the summary for the first 10 countries for the `future` trade data for mass of trade (metric tons), just like the `states` data.    
    
```{r show the data format for summarized trade so far for countries, echo=FALSE}
head(summary_countries_trade_mass_future, n = 10)
```

### Combined Final Trade Data

Now that we have several permutations of trade with *Lycorma*-established states data (both `present` and `future` scenarios) for both geopolitical units (`states` and `countries`) as measured in **value** and **mass** (USD and metric tons, respectively), we can combine some of these data into fewer dataframes to minimize the number of files we are juggling for future analyses. To do this, we merge the **value** and **mass** data for each geopolitical unit dataset.

As we do so, we take the time to create $log_{10}$ transformed versions of the trade **value** and **mass** data. To avoid possible errors for this transformation, we add a constant (`+1`) to data prior to transformation. This process for transformation is held constant for **Impact potential** data.

```{r bind trade data together, message=FALSE}
#STATES
summary_states_trade <- left_join(summary_states_trade_value, summary_states_trade_mass, by = c("destination", "status"), suffix = c("_trade", "_mass")) %>%
    mutate(log10_avg_infected_trade = log10(avg_infected_trade+1), log10_avg_infected_mass = log10(avg_infected_mass+1))

summary_states_trade_future <- left_join(summary_states_trade_value_future, summary_states_trade_mass_future, by = c("destination", "status"), suffix = c("_trade", "_mass")) %>%
    mutate(log10_avg_infected_trade = log10(avg_infected_trade+1), log10_avg_infected_mass = log10(avg_infected_mass+1))

#COUNTRIES

summary_countries_trade <- left_join(summary_countries_trade_value, summary_countries_trade_mass, by = c("destination"), suffix = c("_trade", "_mass")) %>%
    filter(destination != "Western Sahara") %>%
  mutate(log10_avg_infected_trade = log10(avg_infected_trade+1), log10_avg_infected_mass = log10(avg_infected_mass+1))

summary_countries_trade_future <- left_join(summary_countries_trade_value_future, summary_countries_trade_mass_future, by = c("destination"), suffix = c("_trade", "_mass"))  %>%
    filter(destination != "Western Sahara") %>%
  mutate(log10_avg_infected_trade = log10(avg_infected_trade+1), log10_avg_infected_mass = log10(avg_infected_mass+1))
```

> Here is the merged `states` data for `future`:

```{r merged states trade, echo=FALSE}
head(summary_states_trade_future, n = 10)
```

>And here is the merged `countries` data for `future`:

```{r merged countries trade, echo=FALSE}
head(summary_countries_trade_future, n = 10)
```

***

# Impact 

Given the close relationship between cultivation of grapes (`grapes`) and winemaking (`wine`), we evaluate impact potential of *Lycorma* based on data from both.

## Grapes

For `grapes`, two forms of cultivation data were available:

  1. grape production (metric tons)
  2. grape yield per unit area (metric tons/hectare)
  
Although provide both sources of data, we focus more on **production** as a total measure of output per geopolitical unit. 

```{r read in and summarize both grape yield and production, message=FALSE}
data("states_grapes")

#summarize and log transform
summary_states_grapes <- states_grapes %>%
  pivot_wider(names_from = Item, values_from = Value, id_cols = c(geopol_unit, Year)) %>%
  group_by(geopol_unit) %>%
  summarize(avg_yield = mean(grape_yield, na.rm = TRUE), 
            se_yield = sd(grape_yield) / sqrt(n_distinct(Year)),
            avg_prod = mean(grape_production, na.rm = T), se_prod = sd(grape_production) / sqrt(n_distinct(Year))
            ) %>%
  drop_na(.) %>%
  mutate(log10_avg_yield = log10(avg_yield+1), log10_avg_prod = log10(avg_prod+1))

data("countries_grapes")

#first rm Western Sahara, then summarize and log transform, while adding in the step of changing NA's to zeros
summary_countries_grapes <- countries_grapes %>%
  filter(geopol_unit != "Western Sahara") %>%
  pivot_wider(names_from = Item, values_from = Value, id_cols = c(geopol_unit, Year)) %>%
  
  group_by(geopol_unit) %>%
  summarize(avg_yield = mean(grape_yield, na.rm = TRUE), 
            se_yield = sd(grape_yield) / sqrt(n_distinct(Year)),
            avg_prod = mean(grape_production, na.rm = T), se_prod = sd(grape_production) / sqrt(n_distinct(Year))
            ) %>%
  drop_na(.) %>%
  mutate(log10_avg_yield = log10(avg_yield+1), log10_avg_prod = log10(avg_prod+1))

```
  
## Wine

For `wine`, we were able to obtain wine production for both geopolitical unit datasets in units of mass, metric tons. 

```{r read in and summarize wine production, message=FALSE}
data("states_wine")

#summarize and log transform
summary_states_wine <- states_wine %>%
  group_by(geopol_unit) %>%
  summarize(avg_wine = mean(Mass, na.rm = T), se_wine = sd(Mass) / sqrt(n_distinct(Year)), log10_avg_wine = log10(mean(Mass, na.rm = T)+1))

data("countries_wine")

#summarize and log transform, then rm anything that started out as NA or zero
summary_countries_wine <- countries_wine %>%
  group_by(geopol_unit) %>%
  summarize(avg_wine = mean(Mass, na.rm = T), se_wine = sd(Mass) / sqrt(n_distinct(Year)), log10_avg_wine = log10(mean(Mass, na.rm = T)+1)) %>%
  filter(!is.infinite(log10_avg_wine) & !is.nan(log10_avg_wine))

```

# Potentials Summary

Now that all of these data sources have been summarized and tidied, we will merge them together with `left_join()` sequentially, using the `trade` data (**Transport potential**) as the inital dataset that all others are joined to. The resultant datasets contain all permuatations of geopolitical units (`states` and `countries`) and infection scenarios (`present` and `future`).

During this process, we also use the `wine` and `grape` data to produce binary categorical data for each geopolitical unit to indicate whether it is involved directly in that industry. For this, we also manually add the *United States* as a grape producer in the `countries` datasets. Additionally, we manually add `$status` to the `countries` datasets as follows:

  + Native: **China, India, Taiwan, and Vietnam**
  + Established: **Japan, South Korea, and United States**

```{r make full potentials summary datasets, message=FALSE}
#STATES

##PRESENT

#add the binaries
states_summary_present <- summary_states_extracts %>%
  mutate(grapes = ifelse(geopol_unit %in% summary_states_grapes$geopol_unit, yes = "yes", no = "no")) %>%
  mutate(wine = ifelse(geopol_unit %in% summary_states_wine$geopol_unit, yes = "yes", no = "no")) %>%
  mutate(grapes = factor(grapes, levels = c("yes", "no")), wine = factor(wine, levels = c("yes", "no"))) 

#now add in the other data
states_summary_present <- states_summary_present %>%
  left_join(., summary_states_trade, by = c("geopol_unit" = "destination")) %>%   #trade
  left_join(., summary_states_wine, by = "geopol_unit") %>%   #wine
  left_join(., summary_states_grapes, by = "geopol_unit")     #grapes

#deal with NA
states_summary_present[is.na(states_summary_present)] <- 0

#join the new ID's
states_summary_present <- left_join(stateid, states_summary_present, by = "geopol_unit")
#################################
#tidy the order of cols
states_summary_present <- states_summary_present %>%
  select(ID, geopol_unit, status, grand_mean_max, grand_se_max, wine, grapes, avg_wine:log10_avg_prod, everything())

##FUTURE

#add the binaries
states_summary_future <- summary_states_extracts %>%
  mutate(grapes = ifelse(geopol_unit %in% summary_states_grapes$geopol_unit, yes = "yes", no = "no")) %>%
  mutate(wine = ifelse(geopol_unit %in% summary_states_wine$geopol_unit, yes = "yes", no = "no")) %>%
  mutate(grapes = factor(grapes, levels = c("yes", "no")), wine = factor(wine, levels = c("yes", "no"))) 

#now add in the other data
states_summary_future <- states_summary_future %>%
  left_join(., summary_states_trade_future, by = c("geopol_unit" = "destination")) %>%   #trade
  left_join(., summary_states_wine, by = "geopol_unit") %>%   #wine
  left_join(., summary_states_grapes, by = "geopol_unit")     #grapes

#deal with NA
states_summary_future[is.na(states_summary_future)] <- 0

#join the new ID's
states_summary_future <- left_join(stateid, states_summary_future, by = "geopol_unit")

#tidy the order of cols
states_summary_future <- states_summary_future %>%
    select(ID, geopol_unit, status, grand_mean_max, grand_se_max, wine, grapes, avg_wine:log10_avg_prod, everything())



#################################



#COUNTRIES

##PRESENT

#add in the binaries
countries_summary_present <- summary_countries_extracts %>%
  mutate(grapes = ifelse(geopol_unit %in% summary_countries_grapes$geopol_unit, yes = "yes", no = "no")) %>%
  mutate(wine = ifelse(geopol_unit %in% summary_countries_wine$geopol_unit, yes = "yes", no = "no"))

#change U.S. from no to yes for grapes
countries_summary_present$grapes[countries_summary_present$geopol_unit == "United States"] <- "yes"

#The factor fixing step is now here
countries_summary_present <- countries_summary_present %>%
  mutate(grapes = factor(grapes, levels = c("yes", "no")), wine = factor(wine, levels = c("yes", "no"))) 

#now drop NAs for the extracts and add in the other data
countries_summary_present <- countries_summary_present %>%
  drop_na(.) %>%
  left_join(., summary_countries_trade, by = c("geopol_unit" = "destination")) %>%   #trade
  left_join(., summary_countries_wine, by = "geopol_unit") %>%   #wine
  left_join(., summary_countries_grapes, by = "geopol_unit")     #grapes

#deal with NA
countries_summary_present[is.na(countries_summary_present)] <- 0

# add ID to make is simpler for labeling points
countries_summary_present <- countries_summary_present %>%  mutate(ID = geopol_unit)

# add a column for status establishment
countries_summary_present <- countries_summary_present %>%  mutate(status = "not established")
countries_summary_present$status[countries_summary_present$geopol_unit %in% c("China", "India", "Taiwan", "Vietnam")] <- "native"
countries_summary_present$status[countries_summary_present$geopol_unit %in% c("Japan", "South Korea", "United States")] <- "established"

#tidy the order of cols
countries_summary_present <- countries_summary_present %>%
    select(ID, geopol_unit, status, grand_mean_max, grand_se_max, wine, grapes, avg_wine:log10_avg_prod, everything())

##FUTURE

#add in the binaries
countries_summary_future <- summary_countries_extracts %>%
  mutate(grapes = ifelse(geopol_unit %in% summary_countries_grapes$geopol_unit, yes = "yes", no = "no")) %>%
  mutate(wine = ifelse(geopol_unit %in% summary_countries_wine$geopol_unit, yes = "yes", no = "no"))

# NOTE: must This also includes manually changing the **United States** to a grape producer.
#change U.S. from no to yes for grapes
countries_summary_future$grapes[countries_summary_future$geopol_unit == "United States"] <- "yes"

#The factor fixing step is now here
countries_summary_future <- countries_summary_future %>%
  mutate(grapes = factor(grapes, levels = c("yes", "no")), wine = factor(wine, levels = c("yes", "no"))) 

#now drop NAs for the extracts and add in the other data
countries_summary_future <- countries_summary_future %>%
  drop_na(.) %>%
  left_join(., summary_countries_trade_future, by = c("geopol_unit" = "destination")) %>%   #trade
  left_join(., summary_countries_wine, by = "geopol_unit") %>%   #wine
  left_join(., summary_countries_grapes, by = "geopol_unit")     #grapes

#deal with NA
countries_summary_future[is.na(countries_summary_future)] <- 0

# add ID to make is simpler for labeling points
countries_summary_future <- countries_summary_future %>%  mutate(ID = geopol_unit)

# add a column for status establishment
countries_summary_future <- countries_summary_future %>%  mutate(status = "not established")
countries_summary_future$status[countries_summary_future$geopol_unit %in% c("China", "India", "Taiwan", "Vietnam")] <- "native"
countries_summary_future$status[countries_summary_future$geopol_unit %in% c("Japan", "South Korea", "United States")] <- "established"

#tidy the order of cols
countries_summary_future <- countries_summary_future %>%
    select(ID, geopol_unit, status, grand_mean_max, grand_se_max, wine, grapes, avg_wine:log10_avg_prod, everything())

```

***

# Write out data

This section provides the code to save each of the final summary objects to the `/data` directory. They are currently set to not run, as the package **should** ship with the files already in `/data`.

```{r save correlation plot  data, eval=FALSE, message=FALSE}
if(FALSE){
save(states_summary_present, file = file.path(here(), "data", "states_summary_present.rda"))
save(states_summary_future, file = file.path(here(), "data", "states_summary_future.rda"))
save(countries_summary_present, file = file.path(here(), "data", "countries_summary_present.rda"))
save(countries_summary_future, file = file.path(here(), "data", "countries_summary_future.rda"))
}
```
