---
title: "vignette-021-quadrant-plots"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{vignette-021-quadrant-plots}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

new edited version of the correlation plots

To Do:

- V1 : Stack the two figures, ~~flip the axes~~, ~~include the legend~~, ~~remove the quadrant labels~~, ~~make it so that the small white points are visible~~.

  - ~~labeling is wrong on the corrplots for states relative to the infected~~
  - ~~rm observed max suitability in boxplot names~~
  - ~~rm the outliers for boxplots~~
  - ~~scale up the size for the points in the scatter~~
  
- V2: Same as V1, but remove the boxplots.
- V3: Same as V2, but plot colored symbols to indicate the means of yes/no wine.
- V4: Same as V3, but remove the individual points for the states/countries and plot points with SEs on both axes.

```{r setup, results='hide', message=FALSE}
library(slfrsk)
library(tidyverse)
library(here)
library(ggrepel)
library(grid)
library(gridExtra) #only needed if one uses grid.arrange() in plotting
```

read in data
```{r read in summary data}
data("states_summary_present")
data("states_summary_future")
data("countries_summary_present")
data("countries_summary_future")
```

```{r set to run different versions}
# Set these variables to run different versions of the figures.
# Make transport scaled between one and zero (but note that establishment must be uncommented below)
one_zero <- TRUE

# Replace this with max mean quantile_0.50 quantile_0.75 quantile_0.90 to edit which metric to plot for establishment
#to run this, one needs to use different extract data sources further back in the cleaning process (vig01) by selecting a different variable
#estab_to_plot <- "max" # max was used for the quadrant figures

# Need a flag to switch if we want present or future infected states transport
present_transport <- FALSE

# Set what type of multiple correlation you want to calculate
method_cor <- "spearman"

# set whether to look at grape yield or wine production primarily
#impact_prod <- "wine"
impact_prod <- "grapes"

# Plot the native countries?
native <- TRUE

# remove unnecessary "countries" NOTE USA IS REMOVED HERE
#Philippines is also removed, as the grape yield data for it was several orders of magnitude smaller than anything else reported and so close to zero that it caused log transformation scaling issues
rem_usa <- TRUE

# If you want to label both of the countries (or states) with established SLF
label.JPN.KOR <- TRUE
label.us.infected <- TRUE   #PRESENT: adding DE, MD, NJ  FUTURE: adding DE, MD, NJ, OH, WV

# If you want the legend plotted on the states scatter?
# plot_state_legend <- TRUE

```


rescale the trade mass data to be between zero and one or $log_{10}(+1)$ transform the regular version.
```{r rescale states}
# scale the import data

if(one_zero){
  states_summary_present <- states_summary_present %>%
    # transport
    mutate(avg_infected_mass_scaled = log10(avg_infected_mass+1)) %>%
    mutate(avg_infected_mass_scaled = (avg_infected_mass_scaled - min(avg_infected_mass_scaled))) %>%
    mutate(avg_infected_mass_scaled = avg_infected_mass_scaled / max(avg_infected_mass_scaled)) %>%
    # establishment
    mutate(grand_mean_max_scaled = grand_mean_max)
  # mutate(grand_mean_max_scaled = (grand_mean_max - min(grand_mean_max))) %>%
  # mutate(grand_mean_max_scaled = grand_mean_max_scaled / max(grand_mean_max_scaled))
  states_summary_future <- states_summary_future %>%
    # transport
    mutate(avg_infected_mass_scaled = log10(avg_infected_mass+1)) %>%
    mutate(avg_infected_mass_scaled = (avg_infected_mass_scaled - min(avg_infected_mass_scaled))) %>%
    mutate(avg_infected_mass_scaled = avg_infected_mass_scaled / max(avg_infected_mass_scaled)) %>%
    # establishment
    mutate(grand_mean_max_scaled = grand_mean_max)

} else {
  states_summary_present <- states_summary_present %>%
    # transport
    mutate(avg_infected_mass_scaled = log(avg_infected_mass+1)) # %>%
  #mutate(avg_infected_mass_scaled = scale(avg_infected_mass_scaled)) %>%
    # establishment
  #mutate(grand_mean_max_scaled = scale(grand_mean_max))
  states_summary_future <- states_summary_future %>%
    # transport
    mutate(avg_infected_mass_scaled = log(avg_infected_mass+1))

}

```

Modify the native countries established? category?
```{r modify native slf countries}
if(native){
  countries_summary_present$status[countries_summary_present$geopol_unit %in% c("China", "India", "Taiwan", "Vietnam")] <- "native"
  countries_summary_future$status[countries_summary_future$geopol_unit %in% c("China", "India", "Taiwan", "Vietnam")] <- "native"

} else {
  countries_summary_present$status[countries_summary_present$geopol_unit %in% c("China", "India", "Taiwan", "Vietnam")] <- "not established"
  countries_summary_future$status[countries_summary_future$geopol_unit %in% c("China", "India", "Taiwan", "Vietnam")] <- "not established"
  
  }

```

Remove Unnecessary Countries?
```{r rm countries that are not needed}
if(rem_usa){ # TRUE if you want to remove the USA
  rems <- c("Akrotiri and Dhekelia"
            ,"?land"
            ,"American Samoa"
            ,"Bouvet Island"
            ,"British Indian Ocean Territory"
            ,"Caspian Sea"
            ,"Christmas Island"
            ,"Clipperton Island"
            ,"Cocos Islands"
            ,"Falkland Islands"
            ,"Faroe Islands"
            ,"French Southern Territories"
            ,"Heard Island and McDonald Islands"
            ,"Mayotte"
            ,"Northern Mariana Islands"
            ,"Paracel Islands"
            ,"Pitcairn Islands"
            ,"Saint Pierre and Miquelon"
            ,"Tokelau"
            ,"United States Minor Outlying Islands"
            ,"Wallis and Futuna"
            ,"United States"
            ,"Philippines")
} else {
  rems <- c("Akrotiri and Dhekelia"
            ,"?land"
            ,"American Samoa"
            ,"Bouvet Island"
            ,"British Indian Ocean Territory"
            ,"Caspian Sea"
            ,"Christmas Island"
            ,"Clipperton Island"
            ,"Cocos Islands"
            ,"Falkland Islands"
            ,"Faroe Islands"
            ,"French Southern Territories"
            ,"Heard Island and McDonald Islands"
            ,"Mayotte"
            ,"Northern Mariana Islands"
            ,"Paracel Islands"
            ,"Pitcairn Islands"
            ,"Saint Pierre and Miquelon"
            ,"Tokelau"
            ,"United States Minor Outlying Islands"
            ,"Wallis and Futuna"
            ,"Philippines")

}

#do the rm now
countries_summary_present <- countries_summary_present %>% filter(!ID %in% rems)
countries_summary_future <- countries_summary_future %>% filter(!ID %in% rems)

```

Now rescale countries as states
```{r rescale countries}
# scale the import data

if(one_zero){

  countries_summary_present <- countries_summary_present %>%
    # transport
    mutate(avg_infected_mass_scaled = log10(avg_infected_mass+1)) %>%
    mutate(avg_infected_mass_scaled = (avg_infected_mass_scaled - min(avg_infected_mass_scaled))) %>%
    mutate(avg_infected_mass_scaled = avg_infected_mass_scaled / max(avg_infected_mass_scaled)) %>%
    # establishment
    mutate(grand_mean_max_scaled = grand_mean_max)
  # mutate(grand_mean_max_scaled = (grand_mean_max - min(grand_mean_max))) %>%
  # mutate(grand_mean_max_scaled = grand_mean_max_scaled / max(grand_mean_max_scaled))

  countries_summary_future <- countries_summary_future %>%
    # transport
    mutate(avg_infected_mass_scaled = log10(avg_infected_mass+1)) %>%
    mutate(avg_infected_mass_scaled = (avg_infected_mass_scaled - min(avg_infected_mass_scaled))) %>%
    mutate(avg_infected_mass_scaled = avg_infected_mass_scaled / max(avg_infected_mass_scaled)) %>%
    # establishment
    mutate(grand_mean_max_scaled = grand_mean_max)

} else {

  countries_summary_present <- countries_summary_present %>%
    # transport
    mutate(avg_infected_mass_scaled = log(avg_infected_mass+1)) # %>%
  #mutate(avg_infected_mass_scaled = scale(avg_infected_mass_scaled)) %>%
    # establishment
  #mutate(grand_mean_max_scaled = scale(grand_mean_max))
  
  countries_summary_future <- countries_summary_future %>%
    # transport
    mutate(avg_infected_mass_scaled = log(avg_infected_mass+1)) 

}

```

Make a version of the plot that lacks zeros for trade mass
```{r rm zeros}
#rm the trade mass zero countries
countries_summary_nozeros_present <- countries_summary_present %>% filter(log10_avg_infected_mass != 0)
countries_summary_nozeros_future <- countries_summary_future %>% filter(log10_avg_infected_mass != 0)

#rescale and log transform countries with the zeros for trade mass rm
countries_summary_nozeros_present <- countries_summary_nozeros_present %>%
  mutate(avg_infected_mass_scaled = log(avg_infected_mass+1)) %>%
  mutate(avg_infected_mass_scaled = (avg_infected_mass_scaled-min(avg_infected_mass_scaled))) %>%
  mutate(avg_infected_mass_scaled = avg_infected_mass_scaled/max(avg_infected_mass_scaled))
countries_summary_nozeros_future <- countries_summary_nozeros_future %>%
  mutate(avg_infected_mass_scaled = log(avg_infected_mass+1)) %>%
  mutate(avg_infected_mass_scaled = (avg_infected_mass_scaled-min(avg_infected_mass_scaled))) %>%
  mutate(avg_infected_mass_scaled = avg_infected_mass_scaled/max(avg_infected_mass_scaled))

```


Now we have some plotting parameters to make visible up front
```{r params for plotting}
# Choose quadrant intercepts
q_intercepts <- list(transport = 0.5, establishment = 0.5)

hh = .001 #set margins
plots_which <- c(TRUE,FALSE) # Do not change this flips between plotting the states or plotting countries
top_to_plot <- 10 # How many countries and states to highlight for labels?
nudge_JPNKOR.y <- c(.07,.1)
nudge_JPNKOR.x <- c(-.05,0)
nudge_us.infected_present.x <- c(0.00, -0.02,-0.04)  #PRESENT: adding DE, MD, NJ  
nudge_us.infected_present.y <- c(0.10,-0.10,-0.10)
nudge_us.infected_future.x <- c( 0.00,-0.10, 0.02,-0.04,0.00)  #FUTURE: adding OH, DE, MD, NJ, WV
nudge_us.infected_future.y <- c(-0.10, 0.10,-0.17, 0.08,0.15)
#lx <- 0.22
#ly <- 0.25
#hy <- 0.77
#hx <- 0.78
risk_color <- "gray40"
risk_size <- 5

#risk_labels <-data.frame(label = c("low\nrisk","moderate\nrisk","moderate\nrisk","high\nrisk"),
#                        x = c(lx,lx,hx,hx),
#                        y = c(ly,hy,ly,hy),
#                        hjust = c(0,0,0,0),
#                        vjust = c(0,0,0,0))

i <- 1

```


```{r plot figures, warning=FALSE, message=FALSE, echo=FALSE, fig.height= 7.5, fig.width=7}
#for(i in 2){
for (i in 1:length(plots_which)) {
  
  #selects states or countries to plot
  statez = plots_which[i]
  
  ########################
  #DATA SELECTION STEP
  ########################

  #checks whether to use wine or grapes as the impact product to consider
  if(impact_prod == "wine"){
    if (statez) {
      #selects the data to modify for plotting: STATES and chooses the timing
      if(present_transport){
        data_to_plot <- states_summary_present
      } else{
        data_to_plot <- states_summary_future
      }
      
      #calls the exact variables to plot
      data_to_plot <- data_to_plot %>%
        mutate(
          x_to_plot = avg_infected_mass_scaled,
          y_to_plot = grand_mean_max_scaled,
          fill_to_plot = grapes,
          size_to_plot = avg_wine,
          color_to_plot = status
        ) %>%
        arrange(desc(grapes), (size_to_plot)) #so that the grape producers are on top
      
      #order the data that gets labeled based on top_to_plot
      data_to_label <- tail(data_to_plot, top_to_plot)  %>% arrange(desc(size_to_plot))
      #do welch t.test to assess if transport and establishment have the same relationship with grapes
      state_grape_t.test <- list(transport = t.test(avg_infected_mass_scaled~grapes, data = data_to_plot),
                                 establishment = t.test(grand_mean_max_scaled~grapes, data = data_to_plot))
      
      #add infected state labeling if turned on
      if(label.us.infected){
        if(present_transport){  #PRESENT
          data_to_label_us.infected <- data_to_plot %>%
          filter(geopol_unit %in% c("Delaware", "Maryland", "New Jersey")) %>%
          arrange(desc(size_to_plot))
        data_to_label <- bind_rows(data_to_label,data_to_label_us.infected)
        
        } else if(present_transport == FALSE){  #FUTURE
          data_to_label_us.infected <- data_to_plot %>%
          filter(geopol_unit %in% c("Delaware", "Maryland", "New Jersey", "Ohio", "West Virginia")) %>%
          arrange(desc(size_to_plot))
        data_to_label <- bind_rows(data_to_label,data_to_label_us.infected)
        }
      } #end of labeling for infected states
      
    } else {
      #selects the data to modify for plotting: COUNTRIES and selects the timing
      if(present_transport){
        data_to_plot <- countries_summary_present
      } else{
        data_to_plot <- countries_summary_future
      }
      
      data_to_plot <- data_to_plot %>%
        mutate(
          x_to_plot = avg_infected_mass_scaled,
          y_to_plot = grand_mean_max_scaled,
          fill_to_plot = grapes,
          size_to_plot = avg_wine,
          color_to_plot = status
        ) %>%
        arrange(desc(grapes), (size_to_plot)) #so that the grape producers are on top
      data_to_label <-
        tail(data_to_plot, top_to_plot)  %>%
        arrange(desc(size_to_plot)) %>%
        mutate(
          ID = recode(              #add ISO3 codes for the labeled countries (rather than all)
            ID,
            Italy = "ITA",
            France = "FRA",
            Spain = "ESP",
            China = "CHN",
            Argentina = "ARG",
            Chile = "CHL",
            Australia = "AUS",
            `South Africa` = "ZAF",
            Germany = "DEU",
            Portugal = "PRT"
          )
        )
      
      #do welch t.test to assess if transport and establishment have the same relationship with grapes
      country_grape_t.test <- list(transport = t.test(avg_infected_mass_scaled~grapes, data = data_to_plot),
                                   establishment = t.test(grand_mean_max_scaled~grapes, data = data_to_plot))
      if(label.JPN.KOR){
        data_to_label_JPNKOR <- data_to_plot %>%
          filter(geopol_unit %in% c("Japan", "South Korea")) %>%
          mutate(ID = recode(ID,
                             Japan = "JPN",
                             `South Korea` = "KOR")) %>%
          arrange(desc(size_to_plot))
        data_to_label <- bind_rows(data_to_label,data_to_label_JPNKOR)
      }
    }
  } else if(impact_prod == "grapes"){
    if (statez) {
      #selects the data to modify for plotting: STATES and chooses the timing
      if(present_transport){
        data_to_plot <- states_summary_present
      } else{
        data_to_plot <- states_summary_future
      }
      
      data_to_plot <- data_to_plot %>%
        mutate(
          x_to_plot = avg_infected_mass_scaled,
          y_to_plot = grand_mean_max_scaled,
          fill_to_plot = wine,
          size_to_plot = avg_prod, #avg_yield or avg_prod are the two options here for grapes
          color_to_plot = status
        ) %>%
        arrange(desc(grapes), (size_to_plot)) #so that the grape producers are on top
      #change the zeros to one's for plot size
      data_to_plot$size_to_plot[data_to_plot$size_to_plot == 0] <- 1

      data_to_label <-
        tail(data_to_plot, top_to_plot)  %>% arrange(desc(size_to_plot))
      state_grape_t.test <- list(transport = t.test(avg_infected_mass_scaled~grapes, data = data_to_plot),
                                 establishment = t.test(grand_mean_max_scaled~grapes, data = data_to_plot))
      
            #add infected state labeling if turned on
      if(label.us.infected){
        if(present_transport){  #PRESENT
          data_to_label_us.infected <- data_to_plot %>%
          filter(geopol_unit %in% c("Delaware", "Maryland", "New Jersey")) %>%
          arrange(desc(size_to_plot))
        data_to_label <- bind_rows(data_to_label,data_to_label_us.infected)
        
        } else if(present_transport == FALSE){  #FUTURE
          data_to_label_us.infected <- data_to_plot %>%
          filter(geopol_unit %in% c("Delaware", "Maryland", "New Jersey", "Ohio", "West Virginia")) %>%
          arrange(desc(size_to_plot))
        data_to_label <- bind_rows(data_to_label,data_to_label_us.infected)
        }
      } #end of labeling for infected states

    } else {
      #selects the data to modify for plotting: COUNTRIES and selects the timing
      if(present_transport){
        data_to_plot <- countries_summary_present
      } else{
        data_to_plot <- countries_summary_future
      }
      
      data_to_plot <- data_to_plot %>%
        mutate(
          x_to_plot = avg_infected_mass_scaled,
          y_to_plot = grand_mean_max_scaled,
          fill_to_plot = wine,
          size_to_plot = avg_prod, #avg_yield or avg_prod are the two options here for grapes
          color_to_plot = status
        ) %>%
        arrange(desc(grapes), (size_to_plot)) #so that the grape producers are on top
      #change the zeros to one's for plot size
      data_to_plot$size_to_plot[data_to_plot$size_to_plot == 0] <- 1

      data_to_label <-
        tail(data_to_plot, top_to_plot)  %>% arrange(desc(size_to_plot))
      
      data_to_label <- data_to_label %>%
        mutate(
          ID = recode(
            ID,
            Egypt = "EGY",
            Peru = "PER",
            India = "IND",
            Albania = "ALB",
            Iraq = "IRQ",
            Brazil = "BRA",
            Thailand = "THA",
            `South Africa` = "ZAF",
            China = "CHN",
            Armenia = "ARM",
            Italy = "ITA",
            Spain = "ESP",
            France = "FRA",
            Turkey = "TUR",
            Chile = "CHL",
            Argentina = "ARG",
            Australia = "AUS",
            `South Korea` = "KOR",
            Japan = "JPN"
          )
        )
      
      country_grape_t.test <- list(transport = t.test(avg_infected_mass_scaled~grapes, data = data_to_plot),
                                   establishment = t.test(grand_mean_max_scaled~grapes, data = data_to_plot))
      if(label.JPN.KOR){
        data_to_label_JPNKOR <- data_to_plot %>%
          filter(geopol_unit %in% c("Japan", "South Korea")) %>%
          mutate(ID = recode(ID,
                             Japan = "JPN",
                             `South Korea` = "KOR")) %>%
          arrange(desc(size_to_plot))
        data_to_label <- bind_rows(data_to_label,data_to_label_JPNKOR)
      } # end of if label plotting JPN KOR
    }
  }


########################
#PLOT SETTINGS STEP
########################

#SET ALL OF THE NUDGES FOR LABELS
#need to add the conditional for present transport, since nudging only works based on the data
if(present_transport == FALSE){
  if(impact_prod == "wine"){
    #nudging
    if(statez) { #STATES FUTURE WINE
      nudge.x <- c(-.1,-.1,-.1,-.05,-.1,-0.05,-.1,.06,-.01,.05)
      nudge.y <- c(.2,.10,.10,.10,.07,.15,.15,-.07,-.12,-.12)    #CA, WA, NY, PA, OR, GA, OH, MI, VA, NC
      if(label.us.infected){
        nudge.x <- c(nudge.x, nudge_us.infected_future.x)
        nudge.y <- c(nudge.y, nudge_us.infected_future.y)
      }
      ytitle <- "Establishment Potential"
    } else {  #COUNTRIES FUTURE WINE
      nudge.x <- c(-.3,-.27,-.05,-.26,.14,.1,.20,-.21,.08,.2)
      nudge.y <- c(.13, .095, .155, .09, -.11, .13, .15,.06,-.1,-.19)  #ITA, FRA, ESP, CHN, ARG, CHL, AUS, ZAF, DEU, PRT
      if(label.JPN.KOR) {
        nudge.x <- c(nudge.x, nudge_JPNKOR.x) #JPN KOR
        nudge.y <- c(nudge.y, nudge_JPNKOR.y) #JPN KOR
      }
      ytitle <- "Establishment Potential"
    }
  } else if(impact_prod == "grapes"){
    #nudging
    if(statez) { #STATES FUTURE GRAPES
      nudge.x <- c(-.10,-.05,-.10,-.02, .10,-.05, .10, .00,-.05,-.025)   #yield: #CA, PA, WA, MI, NY, MO, OH, OR, GA, NC
      nudge.y <- c(-.10, .10, .08, .08,-.05, .09,-.05,-.10,-.15, .10)   #production: CA, WA, NY, PA, MI, OR, TX, VA, NC, MO
      if(label.us.infected){
        nudge.x <- c(nudge.x, nudge_us.infected_future.x)
        nudge.y <- c(nudge.y, nudge_us.infected_future.y)
      }
      ytitle <- "Establishment Potential"
    } else {  #COUNTRIES FUTURE GRAPES
      # grape production: CHN, ITA, ESP, FRA, TUR, IND, CHL, ARG, ZAF,   AUS
      nudge.x <-       c(-.10,-.25,-.02,-.06, .01,-.05,.200, .10, .075, .145)
      nudge.y <-       c( .05, .11, .12, .09,-.05, .10,.125,-.06,-.1,-.07)
      if(label.JPN.KOR) {
        nudge.x <- c(nudge.x, nudge_JPNKOR.x) #JPN KOR
        nudge.y <- c(nudge.y, nudge_JPNKOR.y) #JPN KOR
      }
      ytitle <- "Establishment Potential"
    }
  }
} else { # grapes above wine below NOTE may not work with labeling JPN and KOR
  if(impact_prod == "wine"){
    if(statez) { #STATES PRESENT WINE
      nudge.x <- c(.09,-.1,-.04,-.05,-.10,-0.12,-.1,.04,-.01,.05)
      nudge.y <- c(-.10,.10,.10,.10,.07,.10,.12,-.07,-.12,-.12)    #CA, WA, NY, PA, OR, GA, OH, MI, VA, NC
      if(label.us.infected){
        nudge.x <- c(nudge.x, nudge_us.infected_present.x)
        nudge.y <- c(nudge.y, nudge_us.infected_present.y)
      }
      ytitle <- "Establishment Potential"
    } else { #COUNTRIES PRESENT WINE
      #nudge.x <- c(0.00,0.00,0.00,0.00,0.00,0.00,0.00,0.00,0.00,0.00)
      #nudge.y <- c(0.00,0.00,0.00,0.00,0.00,0.00,0.00,0.00,0.00,0.00)
      nudge.x <- c(-.03,-.10,-.02,-.03,.14,-.1,-.23,-.21,.08,.2)
      nudge.y <- c(.13, .09, .155, .09, -.11, .13, .15,.09,-.08,-.15)  #ITA, FRA, ESP, CHN, ARG, CHL, AUS, ZAF, DEU, PRT
      if(label.JPN.KOR) {
        nudge.x <- c(nudge.x, nudge_JPNKOR.x) #JPN KOR
        nudge.y <- c(nudge.y, nudge_JPNKOR.y) #JPN KOR
      }
      ytitle <- "Establishment Potential"
    }
  } else if(impact_prod == "grapes"){
    if(statez) { #STATES PRESENT GRAPES
      nudge.y <- c(-0.05, 0.095, -0.10, 0.09, -0.05, 0.085, -0.02, 0.12, -0.10, 0.13)  #CA, WA, NY, PA, MI, OR, TX, VA, NC, MO  
      nudge.x <- c( 0.10, 0.00, -0.02,  0.00, -0.05, 0.00,  0.10, 0.00,  0.02, 0.01)
      if(label.us.infected){
        nudge.x <- c(nudge.x, nudge_us.infected_present.x)
        nudge.y <- c(nudge.y, nudge_us.infected_present.y)
      }
      ytitle <- "Establishment Potential"
    } else { #COUNTRIES PRESENT GRAPES
      nudge.x <- c(-0.01, 0.02, 0.10,-0.04, 0.06,-0.02,-0.07, 0.05,-0.20, 0.11)   #CHN, ITA, ESP, FRA, TUR, IND, CHL, ARG, ZAF, AUS 
      nudge.y <- c( 0.08,-0.08,-0.20, 0.07,-0.02, 0.10, 0.12,-0.09,-0.07,-0.25)
      if(label.JPN.KOR) {
        nudge.x <- c(nudge.x, nudge_JPNKOR.x) #JPN KOR
        nudge.y <- c(nudge.y, nudge_JPNKOR.y) #JPN KOR
      }
      ytitle <- "Establishment Potential"
    }
  }
}
  

  #CREATE THE STATES PLOT
  (states_plot <- ggplot(data = data_to_plot) +
      #geom_text(data = risk_labels, mapping = aes(x = x, y = y,label = label), color = risk_color, size = risk_size) +
      geom_hline(yintercept = q_intercepts$establishment, linetype = "dashed") +
      geom_vline(xintercept = q_intercepts$transport, linetype = "dashed") +
      #  geom_abline(intercept = 0, slope = 1, linetype = "solid") + # add one to one
      geom_rect(mapping = aes(xmin=1.01, xmax=1.2, ymin=.49, ymax=.5), fill = "white") +
      geom_rect(mapping = aes(ymin=1.01, ymax=1.2, xmin=.49, xmax=.51), fill = "white") +
      geom_text_repel(data = data_to_label,
                      aes(x = x_to_plot, y = y_to_plot, label = ID),
                      min.segment.length = 0,
                      direction = "x",
                      nudge_y = nudge.y,
                      nudge_x = nudge.x
      ) +
      geom_point(
        aes(x = x_to_plot, y = y_to_plot, fill = fill_to_plot, size = size_to_plot, color = color_to_plot),
        shape = 21, stroke = 1.3, alpha = 0.75
      ) +
      # geom_text(data = risk_labels, mapping = aes(x = x, y = y,label = label), color = risk_color, size = risk_size) +
      scale_fill_manual(
        values = c("no" = "#ffffff", "yes" = "#C77CFF"),
        name = "Wine Production",# "Wine Impact Potential",
        labels = c("high", "low")
      ) +
      scale_color_manual(
        values = c(
          "red",
          "black",
          "blue"),
        breaks = c(
          "established",
          "not established",
          "native"),
        name = "Regional Status",
        labels = c("invaded","uninvaded", "native")
      ) +
      guides(
        fill = guide_legend(
          order = 1,
          override.aes = list(shape = 22, size = 5, alpha = 1)
        ),
        color = guide_legend(
          order = 3
        ),
        size = guide_legend(    # Adjust size to edit legend grape production
          order = 2,            # circles.
          override.aes = list(size = c(1,3,6))
        )
      ) +
      scale_size_continuous(name = "Grape Production",
                            trans = "log10",
                            range = c(1.5, 6),
                            breaks = c(1.5,2,6),
                            labels = c("low", "moderate", "high")
      ) +
      labs(x = "Transport Potential", y = ytitle) +
      ylim(0, 1.2) + xlim(0, 1.2) +
      theme(
        panel.grid.major = element_line(colour = "#f0f0f0"),
        panel.grid.minor = element_blank(),
        #panel.grid.major = element_blank(),
        #panel.border = element_rect(colour = "black", fill=NA, size=1),
        axis.line = element_line(colour = "black"),
        legend.position = c(0.7, 0.25),
        #legend.position = "bottom",
        panel.background = element_blank(),
        plot.background = element_blank(),
        legend.background = element_blank(),#element_rect(colour = 'black', fill = 'white', linetype='solid'),
        axis.text = element_text(size = rel(1)),
        axis.title.x = element_text(hjust = .4),
        axis.title.y = element_text(hjust = .35),
        legend.title = element_text(face = "plain"),
        legend.key.size = unit(0.2, "cm"),
        legend.key = element_blank(),
        plot.margin = unit(c(hh, -5, hh, hh), units = "line"),
        axis.title = element_text(size = rel(1.3))
      ) +
      scale_x_continuous(breaks = c(0,.5,1),labels = c("low", "moderate", "high")) +
      scale_y_continuous(breaks = c(0,.5,1),labels = c("low", "moderate", "high")) + 
      coord_flip()
  )

  #get the legend
  get_legend <- function(myggplot) {
    tmp <- ggplot_gtable(ggplot_build(myggplot))
    leg <-
      which(sapply(tmp$grobs, function(x)
        x$name) == "guide-box")
    legend <- tmp$grobs[[leg]]
    return(legend)
  }
  #plot all of it
  if(statez) {
    legend_state <- legend <- get_legend(states_plot)
  } else {
    legend_country <- legend <- get_legend(states_plot)
  }


  (states_plot <- states_plot + theme(legend.position = "none"))


  #boxplot: ESTABLISHMENT
  (
    states_box_estab <- ggplot(data_to_plot) +
      geom_boxplot(
        aes(x = fill_to_plot, y = y_to_plot, group = fill_to_plot, fill = fill_to_plot),
        show.legend = FALSE,
        outlier.shape = NA,
        notch = FALSE,
        notchwidth = .25
      ) +
    #labs(x = "Major grape producer", y = "Observed Max Suitability") +
    ylim(0, 1) +
      scale_fill_manual(values = c("yes" = "#C77CFF", "no" = "#ffffff"), name = "") +
      #ggtitle(label = "States Grape Production vs. Average Model Maxes") +
      theme(
        axis.line = element_blank(),
        axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks = element_blank(),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        legend.position = "none",
        panel.background = element_blank(),
        panel.border = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        plot.margin = unit(c(hh, hh, hh, hh), units = "line"),
        plot.background = element_blank()
      ) +
      coord_flip()
  )

  #boxplot: INTRODUCTION
  (
    states_box_intro <- ggplot(data_to_plot) +
      geom_boxplot(
        aes(x = fill_to_plot, y = x_to_plot, group = fill_to_plot, fill = fill_to_plot),
        show.legend = FALSE,
        outlier.shape = NA,
        notch = FALSE,
        notchwidth = .25
      ) +
      #labs(x = "  ", y = "Observed Max Suitability") +
      labs(x = "  ", y = NULL) +
      ylim(0, 1) +
      scale_fill_manual(
        values = c("yes" = "#C77CFF", "no" = "#ffffff"),
        name = ""
      ) +
      #coord_flip() +
      #ggtitle(label = "States Grape Production vs. Average Model Maxes") +
      theme(
        axis.line = element_blank(),
        axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks = element_blank(),
        axis.title.x = element_blank(),
        #axis.title.y=element_blank(),
        legend.position = "none",
        panel.background = element_blank(),
        panel.border = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        plot.background = element_blank(),
        plot.margin = unit(c(hh, hh, hh, hh), units = "line")
      )
  )


  # Insert xbp_grob inside the scatter plot
  #x = x_to_plot,
  #y = y_to_plot
  (
    states_plot_final <- states_plot +
      # annotation_custom(grob = legend, xmin = .465, ymax = .6) +
      #coord_cartesian(clip = "off") +
      annotation_custom(
        grob = ggplotGrob(states_box_estab),
        xmin = 1.01,
        xmax = 1.11,
        ymin = -Inf,
        ymax = 1.05
      ) +
      # Insert ybp_grob inside the scatter plot
      annotation_custom(
        grob = ggplotGrob(states_box_intro),
        xmin = -.1,
        xmax = 1.05,
        ymin = 1,
        ymax = 1.1
      ) +

      annotation_custom(
        grob = rectGrob(gp = gpar(fill = "white", col = "white")),
        xmin = -1,
        xmax = 1.5,
        ymin = 1.1,
        ymax = 1.5
      ) +
      annotation_custom(
        grob = rectGrob(gp = gpar(fill = "white", col = "white")),
        ymin = -1,
        ymax = 1.5,
        xmin = 1.11,
        xmax = 1.5
      )
  )

  g <- ggplotGrob(states_plot_final)
  g$layout$clip[g$layout$name == "panel"] <- "off"
  #g$layout
  g$layout$z[g$layout$name == "panel"] <- 17 # Note that z for panel is 1.  Change it to something bigger.

  grid.newpage()
  grid.draw(g)

  
  #THIS SECTION PLOTS ALL OF IT AGAIN WITH PDF SAVING AND THEN DOES THE CORR
  # present or future
 if(present_transport){
    if (statez) { #states present
      states_grob <- g
      #grid.draw(g)
      data_to_cor <- as.matrix(select(data_to_plot,x_to_plot,y_to_plot,size_to_plot))
      states.impact.model <-
        impact.model <-
        lm(size_to_plot ~ x_to_plot + y_to_plot, data = data.frame(data_to_cor))
      states.impact.cor <- cor.test(impact.model$model$size_to_plot, impact.model$fitted.values, method = method_cor)
    } else {  #countries present
      countries_grob <- g
      #grid.draw(g)
      data_to_cor <- as.matrix(select(data_to_plot,x_to_plot,y_to_plot,size_to_plot))
      countries.impact.model <-
        impact.model <-
        lm(size_to_plot ~ x_to_plot + y_to_plot, data = data.frame(data_to_cor))
      countries.impact.cor <- cor.test(impact.model$model$size_to_plot, impact.model$fitted.values, method = method_cor)

  } 
   }  else {    #future cases

    if (statez) { #states future
      states_grob <- g
      grid.draw(g)
      data_to_cor <- as.matrix(select(data_to_plot,x_to_plot,y_to_plot,size_to_plot))
      states.impact.model <-
        impact.model <-
        lm(size_to_plot ~ x_to_plot + y_to_plot, data = data.frame(data_to_cor))
      states.impact.cor <- cor.test(impact.model$model$size_to_plot, impact.model$fitted.values, method = method_cor)
    } else{ #countries future
      countries_grob <- g
      grid.draw(g)
      data_to_cor <- as.matrix(select(data_to_plot,x_to_plot,y_to_plot,size_to_plot))
      countries.impact.model <-
        impact.model <-
        lm(size_to_plot ~ x_to_plot + y_to_plot, data = data.frame(data_to_cor))
      countries.impact.cor <- cor.test(impact.model$model$size_to_plot, impact.model$fitted.values, method = method_cor)
      } 
            }

  #plot the states+legend and save
if (statez) {
  if(present_transport){
    pdf(file.path(here::here(),"vignettes", paste0("states_combined_present_max_",impact_prod,".pdf")),width = 7.5, height = 7)

} else{
    pdf(file.path(here::here(),"vignettes", paste0("states_combined_future_max_",impact_prod,".pdf")),width = 7.5, height = 7)
      
}
      grid.arrange(states_grob, legend_state, nrow=1, ncol = 2, widths = c(4,2), heights = 1)
    dev.off()
} else{
  #plot the countries+legend
if(present_transport){
    pdf(file.path(here::here(),"vignettes", paste0("countries_combined_present_max_",impact_prod,".pdf")),width = 7.5, height = 7)

} else{
    pdf(file.path(here::here(),"vignettes", paste0("countries_combined_future_max_",impact_prod,".pdf")),width = 7.5, height = 7)
      
}
      grid.arrange(countries_grob, legend_country, nrow=1, ncol = 2, widths = c(4,2), heights = 1)
    dev.off()

}
  
} # end of the 2 loop for loop for plotting states and countries




  lay <- rbind(c(1,1,1,NA,NA,NA),
               c(1,1,1,NA,NA,NA),
               c(1,1,1,NA,3,3),
               c(NA,NA,NA,NA,3,3),
               c(2,2,2,NA,3,3),
               c(2,2,2,NA,NA,NA),
               c(2,2,2,NA,NA,NA)
  )
  
  grid.arrange(states_grob, countries_grob, legend_country, nrow = 7, ncol = 6, layout_matrix = lay, heights = c(1,1,1, 0.22, 1,1,1), widths = c(0.8,0.8,0.8,0.01,0.7,0.7))
```


