---
title: "vignette-020-quadrant-plots"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{vignette-020-quadrant-plots}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This is the standard version of the correlation plot code!

```{r setup}
library(slfRisk)
library(tidyverse)
library(here)
library(ggrepel)
library(grid)
#library(gridExtra) #only needed if one uses grid.arrange() in plotting
```

read in data
```{r read in summary data}
data("states_summary_present")
data("states_summary_future")
data("countries_summary_present")
data("countries_summary_future")
```

```{r set to run different versions}
# Set these variables to run different versions of the figures.

# Make transport scaled between one and zero (but note that establishment must be uncommented below)
one_zero <- TRUE

# Replace this with max mean quantile_0.50 quantile_0.75 quantile_0.90 to edit which metric to plot for establishment
estab_to_plot <- "max" # max was used for the quadrant figures

# Choose quadrant intercepts
q_intercepts <- list(transport = 0.5, establishment = 0.5) 

# Need a flag to switch if we want present or future infected states transport
present_transport <- FALSE

# Set what type of multiple correlation you want to calculate
method_cor <- "spearman"

# set whether to look at grape yield or wine production primarily
#impact_prod <- "wine"
impact_prod <- "grapes"

# Set the version number here for the pdfs that are being produced
vers <- paste0("4_4_prod_", impact_prod)


# Plot the native countries?
native <- TRUE

# If you want to label both of the countries with established SLF
label.JPN.KOR <- TRUE

# If you want the legend plotted on the states scatter?
# plot_state_legend <- TRUE

# remove unnecessary "countries" NOTE USA IS REMOVED HERE
#Philippines is also removed, as the grape yield data for it was several orders of magnitude smaller than anything else reported and so close to zero that it caused log transformation scaling issues
rem_usa <- TRUE

```


rescale the trade mass data to be between zero and one or $log_{10}(+1)$ transform the regular version.
```{r rescale states}
# scale the import data

if(one_zero){
  states_summary_present <- states_summary_present %>%
    # transport
    mutate(avg_infected_mass_scaled = log10(avg_infected_mass+1)) %>%
    mutate(avg_infected_mass_scaled = (avg_infected_mass_scaled - min(avg_infected_mass_scaled))) %>%
    mutate(avg_infected_mass_scaled = avg_infected_mass_scaled / max(avg_infected_mass_scaled)) %>%
    # establishment
    mutate(grand_mean_max_scaled = grand_mean_max)
  # mutate(grand_mean_max_scaled = (grand_mean_max - min(grand_mean_max))) %>%
  # mutate(grand_mean_max_scaled = grand_mean_max_scaled / max(grand_mean_max_scaled))
  states_summary_future <- states_summary_future %>%
    # transport
    mutate(avg_infected_mass_scaled = log10(avg_infected_mass+1)) %>%
    mutate(avg_infected_mass_scaled = (avg_infected_mass_scaled - min(avg_infected_mass_scaled))) %>%
    mutate(avg_infected_mass_scaled = avg_infected_mass_scaled / max(avg_infected_mass_scaled)) %>%
    # establishment
    mutate(grand_mean_max_scaled = grand_mean_max)

} else {
  states_summary_present <- states_summary_present %>%
    # transport
    mutate(avg_infected_mass_scaled = log(avg_infected_mass+1)) # %>%
  #mutate(avg_infected_mass_scaled = scale(avg_infected_mass_scaled)) %>%
    # establishment
  #mutate(grand_mean_max_scaled = scale(grand_mean_max))
  states_summary_future <- states_summary_future %>%
    # transport
    mutate(avg_infected_mass_scaled = log(avg_infected_mass+1))

}

```

Modify the native countries established? category?
```{r modify native slf countries}
if(native){
  countries_summary_present$status[countries_summary_present$geopol_unit %in% c("China", "India", "Taiwan", "Vietnam")] <- "native"
  countries_summary_future$status[countries_summary_future$geopol_unit %in% c("China", "India", "Taiwan", "Vietnam")] <- "native"

} else {
  countries_summary_present$status[countries_summary_present$geopol_unit %in% c("China", "India", "Taiwan", "Vietnam")] <- "not established"
  countries_summary_future$status[countries_summary_future$geopol_unit %in% c("China", "India", "Taiwan", "Vietnam")] <- "not established"
  
  }

```

Remove Unnecessary Countries?
```{r rm countries that are not needed}
if(rem_usa){ # TRUE if you want to remove the USA
  rems <- c("Akrotiri and Dhekelia"
            ,"?land"
            ,"American Samoa"
            ,"Bouvet Island"
            ,"British Indian Ocean Territory"
            ,"Caspian Sea"
            ,"Christmas Island"
            ,"Clipperton Island"
            ,"Cocos Islands"
            ,"Falkland Islands"
            ,"Faroe Islands"
            ,"French Southern Territories"
            ,"Heard Island and McDonald Islands"
            ,"Mayotte"
            ,"Northern Mariana Islands"
            ,"Paracel Islands"
            ,"Pitcairn Islands"
            ,"Saint Pierre and Miquelon"
            ,"Tokelau"
            ,"United States Minor Outlying Islands"
            ,"Wallis and Futuna"
            ,"United States"
            ,"Philippines")
} else {
  rems <- c("Akrotiri and Dhekelia"
            ,"?land"
            ,"American Samoa"
            ,"Bouvet Island"
            ,"British Indian Ocean Territory"
            ,"Caspian Sea"
            ,"Christmas Island"
            ,"Clipperton Island"
            ,"Cocos Islands"
            ,"Falkland Islands"
            ,"Faroe Islands"
            ,"French Southern Territories"
            ,"Heard Island and McDonald Islands"
            ,"Mayotte"
            ,"Northern Mariana Islands"
            ,"Paracel Islands"
            ,"Pitcairn Islands"
            ,"Saint Pierre and Miquelon"
            ,"Tokelau"
            ,"United States Minor Outlying Islands"
            ,"Wallis and Futuna"
            ,"Philippines")

}

#do the rm now
countries_summary_present <- countries_summary_present %>% filter(!ID %in% rems)
countries_summary_future <- countries_summary_future %>% filter(!ID %in% rems)

```

Now rescale countries as states
```{r rescale countries}
# scale the import data

if(one_zero){

  countries_summary_present <- countries_summary_present %>%
    # transport
    mutate(avg_infected_mass_scaled = log10(avg_infected_mass+1)) %>%
    mutate(avg_infected_mass_scaled = (avg_infected_mass_scaled - min(avg_infected_mass_scaled))) %>%
    mutate(avg_infected_mass_scaled = avg_infected_mass_scaled / max(avg_infected_mass_scaled)) %>%
    # establishment
    mutate(grand_mean_max_scaled = grand_mean_max)
  # mutate(grand_mean_max_scaled = (grand_mean_max - min(grand_mean_max))) %>%
  # mutate(grand_mean_max_scaled = grand_mean_max_scaled / max(grand_mean_max_scaled))

  countries_summary_future <- countries_summary_future %>%
    # transport
    mutate(avg_infected_mass_scaled = log10(avg_infected_mass+1)) %>%
    mutate(avg_infected_mass_scaled = (avg_infected_mass_scaled - min(avg_infected_mass_scaled))) %>%
    mutate(avg_infected_mass_scaled = avg_infected_mass_scaled / max(avg_infected_mass_scaled)) %>%
    # establishment
    mutate(grand_mean_max_scaled = grand_mean_max)

} else {

  countries_summary_present <- countries_summary_present %>%
    # transport
    mutate(avg_infected_mass_scaled = log(avg_infected_mass+1)) # %>%
  #mutate(avg_infected_mass_scaled = scale(avg_infected_mass_scaled)) %>%
    # establishment
  #mutate(grand_mean_max_scaled = scale(grand_mean_max))
  
  countries_summary_future <- countries_summary_future %>%
    # transport
    mutate(avg_infected_mass_scaled = log(avg_infected_mass+1)) 

}

```

Make a version of the plot that lacks zeros for trade mass
```{r}
#rm the trade mass zero countries
countries_summary_nozeros_present <- countries_summary_present %>% filter(log10_avg_infected_mass != 0)
countries_summary_nozeros_future <- countries_summary_future %>% filter(log10_avg_infected_mass != 0)

#rescale and log transform countries with the zeros for trade mass rm
countries_summary_nozeros_present <- countries_summary_nozeros_present %>%
  mutate(avg_infected_mass_scaled = log(avg_infected_mass+1)) %>%
  mutate(avg_infected_mass_scaled = (avg_infected_mass_scaled-min(avg_infected_mass_scaled))) %>%
  mutate(avg_infected_mass_scaled = avg_infected_mass_scaled/max(avg_infected_mass_scaled))
countries_summary_nozeros_future <- countries_summary_nozeros_future %>%
  mutate(avg_infected_mass_scaled = log(avg_infected_mass+1)) %>%
  mutate(avg_infected_mass_scaled = (avg_infected_mass_scaled-min(avg_infected_mass_scaled))) %>%
  mutate(avg_infected_mass_scaled = avg_infected_mass_scaled/max(avg_infected_mass_scaled))

```


Now we have some plotting parameters to make visible up front
```{r params for plotting}
# Choose quadrant intercepts
q_intercepts <- list(transport = 0.5, establishment = 0.5)

hh = .01 #set margins
plots_which <- c(TRUE,FALSE) # Do not change this flips between potting the states or plotting countries
top_to_plot <- 10 # How many countries and states to highlight for labels?
nudge_JPNKOR.y <- c(.1,.1)
nudge_JPNKOR.x <- c(-.17,0)
lx <- 0.22
ly <- 0.25
hy <- 0.77
hx <- 0.78
risk_color <- "gray40"
risk_size <- 5

risk_labels <-data.frame(label = c("low\nrisk","moderate\nrisk","moderate\nrisk","high\nrisk"),
                         x = c(lx,lx,hx,hx),
                         y = c(ly,hy,ly,hy),
                         hjust = c(0,0,0,0),
                         vjust = c(0,0,0,0))

i <- 1

```

And the actual plots...
```{r plot figures?}
for (i in 1:length(plots_which)) {
  # for (i in 1) {
  #  for (i in 2) {
  statez = plots_which[i]

  if(impact_prod == "wine"){
    if (statez) {
      data_to_plot <- states_summary_present
      data_to_plot <- data_to_plot %>%
        mutate(
          x_to_plot = avg_infected_mass_scaled,
          y_to_plot = grand_mean_max_scaled,
          fill_to_plot = grapes,
          size_to_plot = avg_wine,
          color_to_plot = status
        ) %>%
        arrange(desc(grapes), (size_to_plot)) #so that the grape producers are on top
      data_to_label <-
        tail(data_to_plot, top_to_plot)  %>% arrange(desc(size_to_plot))
      state_grape_t.test <- list(transport = t.test(avg_infected_mass_scaled~grapes, data = data_to_plot),
                                 establishment = t.test(grand_mean_max_scaled~grapes, data = data_to_plot))
    } else {
      data_to_plot <- countries_summary_present
      data_to_plot <- data_to_plot %>%
        mutate(
          x_to_plot = avg_infected_mass_scaled,
          y_to_plot = grand_mean_max_scaled,
          fill_to_plot = grapes,
          size_to_plot = avg_wine,
          color_to_plot = status
        ) %>%
        arrange(desc(grapes), (size_to_plot)) #so that the grape producers are on top
      data_to_label <-
        tail(data_to_plot, top_to_plot)  %>%
        arrange(desc(size_to_plot)) %>%
        mutate(
          ID = recode(
            ID,
            Italy = "ITA",
            France = "FRA",
            Spain = "ESP",
            China = "CHN",
            Argentina = "ARG",
            Chile = "CHL",
            Australia = "AUS",
            `South Africa` = "ZAF",
            Germany = "DEU",
            Portugal = "PRT"
          )
        )

      country_grape_t.test <- list(transport = t.test(avg_infected_mass_scaled~grapes, data = data_to_plot),
                                   establishment = t.test(grand_mean_max_scaled~grapes, data = data_to_plot))
      if(label.JPN.KOR){
        data_to_label_JPNKOR <- data_to_plot %>%
          filter(geopol_unit %in% c("Japan", "South Korea")) %>%
          mutate(ID = recode(ID,
                             Japan = "JPN",
                             `South Korea` = "KOR")) %>%
          arrange(desc(size_to_plot))
        data_to_label <- bind_rows(data_to_label,data_to_label_JPNKOR)
      }
    }
  } else if(impact_prod == "grapes"){
    if (statez) {
      data_to_plot <- states_summary_present
      data_to_plot <- data_to_plot %>%
        mutate(
          x_to_plot = avg_infected_mass_scaled,
          y_to_plot = grand_mean_max_scaled,
          fill_to_plot = wine,
          size_to_plot = avg_prod, #avg_yield or avg_prod are the two options here for grapes
          color_to_plot = status
        ) %>%
        arrange(desc(grapes), (size_to_plot)) #so that the grape producers are on top
      #change the zeros to one's for plot size
      data_to_plot$size_to_plot[data_to_plot$size_to_plot == 0] <- 1

      data_to_label <-
        tail(data_to_plot, top_to_plot)  %>% arrange(desc(size_to_plot))
      state_grape_t.test <- list(transport = t.test(avg_infected_mass_scaled~grapes, data = data_to_plot),
                                 establishment = t.test(grand_mean_max_scaled~grapes, data = data_to_plot))
    } else {
      data_to_plot <- countries_summary_present
      data_to_plot <- data_to_plot %>%
        mutate(
          x_to_plot = avg_infected_mass_scaled,
          y_to_plot = grand_mean_max_scaled,
          fill_to_plot = wine,
          size_to_plot = avg_prod, #avg_yield or avg_prod are the two options here for grapes
          color_to_plot = status
        ) %>%
        arrange(desc(grapes), (size_to_plot)) #so that the grape producers are on top
      #change the zeros to one's for plot size
      data_to_plot$size_to_plot[data_to_plot$size_to_plot == 0] <- 1

      data_to_label <-
        tail(data_to_plot, top_to_plot)  %>% arrange(desc(size_to_plot))
      data_to_label <- data_to_label %>%
        mutate(
          ID = recode(
            ID,
            Egypt = "EGY",
            Peru = "PER",
            India = "IND",
            Albania = "ALB",
            Iraq = "IRQ",
            Brazil = "BRA",
            Thailand = "THA",
            `South Africa` = "ZAF",
            China = "CHN",
            Armenia = "ARM",
            Italy = "ITA",
            Spain = "ESP",
            France = "FRA",
            Turkey = "TUR",
            Chile = "CHL",
            Argentina = "ARG",
            Australia = "AUS",
            `South Korea` = "KOR",
            Japan = "JPN"
          )
        )
      country_grape_t.test <- list(transport = t.test(avg_infected_mass_scaled~grapes, data = data_to_plot),
                                   establishment = t.test(grand_mean_max_scaled~grapes, data = data_to_plot))
      if(label.JPN.KOR){
        data_to_label_JPNKOR <- data_to_plot %>%
          filter(geopol_unit %in% c("Japan", "South Korea")) %>%
          mutate(ID = recode(ID,
                             Japan = "JPN",
                             `South Korea` = "KOR")) %>%
          arrange(desc(size_to_plot))
        data_to_label <- bind_rows(data_to_label,data_to_label_JPNKOR)
      } # end of if label plotting JPN KOR
    }
  }



  #plot all of it
  #need to add the conditional for present transport, since nudging only works based on the data
  if(present_transport == FALSE){
    if(impact_prod == "wine"){
      #nudging
      if(statez) {
        nudge.y <- c(.2,.10,.10,.10,.07,.15,.15,-.07,-.12,-.12)    #CA, WA, NY, PA, OR, GA, OH, MI, VA, NC
        nudge.x <- c(-.1,-.1,-.1,-.05,-.1,-0.05,-.1,.06,-.01,.05)
        ytitle <- "Establishment Potential"
      } else {
        nudge.y <- c(.13, .095, .155, .09, -.11, .13, .15,.06,-.1,-.19)  #ITA, FRA, ESP, CHN, ARG, CHL, AUS, ZAF, DEU, PRT
        nudge.x <- c(-.3,-.27,-.05,-.26,.14,.1,.20,-.21,.08,.2)
        if(label.JPN.KOR) {
          nudge.x <- c(nudge.x, nudge_JPNKOR.x) #JPN KOR
          nudge.y <- c(nudge.y, nudge_JPNKOR.y) #JPN KOR
        }
        ytitle <- "Establishment Potential"
      }
    } else if(impact_prod == "grapes"){
      #nudging
      if(statez) {
        nudge.y <- c( .20, .10, .10, .10,-.05, .10,-.05,-.07,-.12, .10)   #production: CA, WA, NY, PA, MI, OR, TX, VA, NC, MO
        nudge.x <- c(-.10,-.05,-.22,-.05, .10,-.05, .10,.025, .05,-.025)   #yield: #CA, PA, WA, MI, NY, MO, OH, OR, GA, NC
        ytitle <- "Establishment Potential"
      } else {
        # grape production: CHN, ITA, ESP, FRA, TUR, IND, CHL, ARG, ZAF,   AUS
        nudge.y <-       c( .10, .11, .12, .09,-.05, .10,.125,-.06,-.1,-.07)
        nudge.x <-       c(-.05,-.25,-.02,-.06, .01,-.05,.200, .10, .075, .145)
        if(label.JPN.KOR) {
          nudge.x <- c(nudge.x, nudge_JPNKOR.x) #JPN KOR
          nudge.y <- c(nudge.y, nudge_JPNKOR.y) #JPN KOR
        }
        ytitle <- "Establishment Potential"
      }
    }

  } else { # grapes above wine below NOTE will not work with labeling JPN and KOR
    if(statez) {
      nudge.y <- c(-.10,.10,.10,.10,.07,.10,.12,-.07,-.12,-.12)    #CA, WA, NY, PA, OR, GA, OH, MI, VA, NC
      nudge.x <- c(.09,-.1,-.04,-.05,-.10,-0.12,-.1,.04,-.01,.05)
      ytitle <- "Establishment Potential"
    } else {
      nudge.y <- c(.13, .09, .155, .09, -.11, .13, .15,.09,-.08,-.15)  #ITA, FRA, ESP, CHN, ARG, CHL, AUS, ZAF, DEU, PRT
      nudge.x <- c(-.03,-.10,-.02,-.03,.14,-.1,-.23,-.21,.08,.2)
      ytitle <- "Establishment Potential"
    }
  }

  (states_plot <- ggplot(data = data_to_plot) +
      geom_text(data = risk_labels,
               mapping = aes(x = x, y = y,label = label),
               color = risk_color,
               size = risk_size) +
      #geom_segment(aes(x = 0, xend = 1, y = .5, yend = .5), linetype = "dashed") +
      #geom_segment(aes(y = 0, yend = 1, x = .5, xend = .5), linetype = "dashed") +
      geom_hline(yintercept = q_intercepts$establishment, linetype = "dashed") +
      geom_vline(xintercept = q_intercepts$transport, linetype = "dashed") +
      #  geom_abline(intercept = 0, slope = 1, linetype = "solid") + # add one to one
      geom_rect(mapping = aes(xmin=1.01, xmax=1.2, ymin=.49, ymax=.5), fill = "white") +
      geom_rect(mapping = aes(ymin=1.01, ymax=1.2, xmin=.49, xmax=.51), fill = "white") +
      geom_text_repel(data = data_to_label,
                      aes(x = x_to_plot, y = y_to_plot, label = ID),
                      min.segment.length = 0,
                      direction = "x",
                      nudge_y = nudge.y,
                      nudge_x = nudge.x
      ) +
      geom_point(
        aes(x = x_to_plot, y = y_to_plot,
            fill = fill_to_plot, size = size_to_plot, color = color_to_plot),
        shape = 21, stroke = 1.3, alpha = 0.75
      ) +
      # geom_text(data = risk_labels,
      #           mapping = aes(x = x, y = y,label = label),
      #           color = risk_color,
      #           size = risk_size) +
      scale_fill_manual(
        values = c("no" = "#ffffff", "yes" = "#C77CFF"),
        name = "Wine Production",# "Wine Impact Potential",
        labels = c("high", "low")
      ) +
      scale_color_manual(
        values = c(
          "established" = "red",
          "not established" = "black",
          "native" = "blue"),
        name = "Regional Status",
        labels = c("invaded","native", "uninvaded")
      ) +
      guides(
        fill = guide_legend(
          order = 1,
          override.aes = list(
            shape = 22,
            size = 5,
            alpha = 1
          )
        ),
        color = guide_legend(
          order = 3
        ),
        size = guide_legend(    # Adjust size to edit legend grape production
          order = 2,            # circles.
          override.aes = list(
            size = c(1,3,6)
          )
        )
      ) +
      scale_size_continuous(name = "Grape Production",
                            trans = "log10",
                            range = c(1, 6),
                            breaks = c(1,2,6),
                            labels = c("low", "moderate", "high")
      ) +
      labs(x = "Transport Potential", y = ytitle) +
      ylim(0, 1.2) + xlim(0, 1.2) +
      theme(
        panel.grid.major = element_line(colour = "#f0f0f0"),
        panel.grid.minor = element_blank(),
        #panel.grid.major = element_blank(),
        #panel.border = element_rect(colour = "black", fill=NA, size=1),
        axis.line = element_line(colour = "black"),
        legend.position = c(0.7, 0.25),
        #legend.position = "bottom",
        panel.background = element_blank(),
        plot.background = element_blank(),
        legend.background = element_blank(),#element_rect(colour = 'black', fill = 'white', linetype='solid'),
        axis.text = element_text(size = rel(1)),
        axis.title.x = element_text(hjust = .4),
        axis.title.y = element_text(hjust = .35),
        legend.title = element_text(face = "plain"),
        legend.key.size = unit(0.2, "cm"),
        legend.key = element_blank(),
        plot.margin = unit(c(hh, hh, hh, hh), units = "line"),
        axis.title = element_text(size = rel(1.3))
      ) +
      scale_x_continuous(breaks = c(0,.5,1),labels = c("low", "moderate", "high")) +
      scale_y_continuous(breaks = c(0,.5,1),labels = c("low", "moderate", "high"))
      #+ coord_flip()
  )

  #get the legend
  get_legend <- function(myggplot) {
    tmp <- ggplot_gtable(ggplot_build(myggplot))
    leg <-
      which(sapply(tmp$grobs, function(x)
        x$name) == "guide-box")
    legend <- tmp$grobs[[leg]]
    return(legend)
  }
  #plot all of it
  if(statez) {
    legend_state <- legend <- get_legend(states_plot)
  } else {
    legend_country <- legend <- get_legend(states_plot)
  }


  (states_plot <- states_plot + theme(legend.position = "none"))


  #boxplot
  (
    states_box_estab <- ggplot(data_to_plot) +
      geom_boxplot(
        aes(
          x = fill_to_plot,
          y = y_to_plot,
          group = fill_to_plot,
          fill = fill_to_plot
        ),
        show.legend = FALSE,
        #outlier.size = 0,
        notch = FALSE,
        notchwidth = .25
      ) +
      #geom_dotplot(aes(x = fill_to_plot, y = y_to_plot, group = fill_to_plot, fill = fill_to_plot),
      #             binaxis='y',
      #             stackdir='center',
      #             dotsize=1) +
      # geom_point(aes(x = fill_to_plot,
      #                y = y_to_plot,
      #                fill = fill_to_plot,
      #                size = log10(size_to_plot),
      #                color = color_to_plot),
      #            shape = 21,
      #            stroke = 1.3) +
    #labs(x = "Major grape producer", y = "Observed Max Suitability") +
    ylim(0, 1) +
      scale_fill_manual(
        values = c("yes" = "#C77CFF", "no" = "#ffffff"),
        name = ""
      ) +
      #ggtitle(label = "States Grape Production vs. Average Model Maxes") +
      theme(
        axis.line = element_blank(),
        axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks = element_blank(),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        legend.position = "none",
        panel.background = element_blank(),
        panel.border = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        plot.margin = unit(c(hh, hh, hh, hh), units = "line"),
        plot.background = element_blank()
      )
  )

  #ggplot(data = data_to_plot) + geom_density(mapping = aes(x_to_plot, fill = fill_to_plot))
  #boxplot
  (
    states_box_intro <- ggplot(data_to_plot) +
      geom_boxplot(
        aes(
          x = fill_to_plot,
          y = x_to_plot,
          group = fill_to_plot,
          fill = fill_to_plot
        ),
        show.legend = FALSE,
        #outlier.size = 0,
        notch = FALSE,
        notchwidth = .25
      ) +
      # geom_point(aes(x = fill_to_plot,
      #                y = x_to_plot,
      #                fill = fill_to_plot,
      #                size = log10(size_to_plot),
      #                color = color_to_plot),
      #            shape = 21,
      #            stroke = 1.3) +
      #geom_density(aes(x_to_plot, fill = fill_to_plot), show.legend = F, alpha = 0.6) +
      labs(x = "  ", y = "Observed Max Suitability") +
      ylim(0, 1) +
      scale_fill_manual(
        values = c("yes" = "#C77CFF", "no" = "#ffffff"),
        name = ""
      ) +
      coord_flip() +
      #ggtitle(label = "States Grape Production vs. Average Model Maxes") +
      theme(
        axis.line = element_blank(),
        axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks = element_blank(),
        axis.title.x = element_blank(),
        #axis.title.y=element_blank(),
        legend.position = "none",
        panel.background = element_blank(),
        panel.border = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        plot.background = element_blank(),
        plot.margin = unit(c(hh, hh, hh, hh), units = "line")
      )
  )


  #arrange it in a nice combined grid

  #grid.arrange(states_box_intro, grid.text(""), states_plot, states_box_estab, nrow = 2, ncol = 2, widths = c(5, 1), heights = c(1,5))



  # Insert xbp_grob inside the scatter plot
  #x = x_to_plot,
  #y = y_to_plot
  (
    states_plot_final <- states_plot +
      # annotation_custom(grob = legend, xmin = .465, ymax = .6) +
      #coord_cartesian(clip = "off") +
      annotation_custom(
        grob = ggplotGrob(states_box_estab),
        xmin = 1.01,
        xmax = 1.11,
        ymin = -Inf,
        ymax = 1.05
      ) +
      # Insert ybp_grob inside the scatter plot
      annotation_custom(
        grob = ggplotGrob(states_box_intro),
        xmin = -.1,
        xmax = 1.05,
        ymin = 1,
        ymax = 1.1
      ) +

      annotation_custom(
        grob = rectGrob(gp = gpar(fill = "white", col = "white")),
        xmin = -1,
        xmax = 1.5,
        ymin = 1.1,
        ymax = 1.5
      ) +
      annotation_custom(
        grob = rectGrob(gp = gpar(fill = "white", col = "white")),
        ymin = -1,
        ymax = 1.5,
        xmin = 1.11,
        xmax = 1.5
      )
  )

  g <- ggplotGrob(states_plot_final)
  g$layout$clip[g$layout$name == "panel"] <- "off"
  #g$layout
  g$layout$z[g$layout$name == "panel"] <-
    17 # Note that z for panel is 1.  Change it to something bigger.

  grid.newpage()
  grid.draw(g)

  # present or future
  if(present_transport){

    if (statez) {
      states_grob <- g
      #pdf(paste0("./submission/Global_MAXENT/figures/states_combined_",estab_to_plot,"_v",vers,".pdf"), width = 7.5, height = 7)
      grid.draw(g)
      #dev.off()
      data_to_cor <- as.matrix(select(data_to_plot,x_to_plot,y_to_plot,size_to_plot))
      states.impact.model <-
        impact.model <-
        lm(size_to_plot ~ x_to_plot + y_to_plot, data = data.frame(data_to_cor))
      states.impact.cor <- cor.test(impact.model$model$size_to_plot,
                                    impact.model$fitted.values,
                                    method = method_cor)
    } else {
      countries_grob <- g
      #pdf(paste0("./submission/Global_MAXENT/figures/countries_combined_",estab_to_plot,"_v",vers,".pdf"), width = 7.5, height = 7)
      grid.draw(g)
      #dev.off()
      data_to_cor <- as.matrix(select(data_to_plot,x_to_plot,y_to_plot,size_to_plot))
      countries.impact.model <-
        impact.model <-
        lm(size_to_plot ~ x_to_plot + y_to_plot, data = data.frame(data_to_cor))
      countries.impact.cor <- cor.test(impact.model$model$size_to_plot,
                                       impact.model$fitted.values,
                                       method = method_cor)
    }
  } else {


    if (statez) {
      states_grob <- g
      #pdf(paste0("./submission/Global_MAXENT/figures/states_combined_future_",estab_to_plot,"_v",vers,".pdf"), width = 7.5, height = 7)
      grid.draw(g)
      #dev.off()
      data_to_cor <- as.matrix(select(data_to_plot,x_to_plot,y_to_plot,size_to_plot))
      states.impact.model <-
        impact.model <-
        lm(size_to_plot ~ x_to_plot + y_to_plot, data = data.frame(data_to_cor))
      states.impact.cor <- cor.test(impact.model$model$size_to_plot,
                                    impact.model$fitted.values,
                                    method = method_cor)
    } else {
      countries_grob <- g
      #pdf(paste0("./submission/Global_MAXENT/figures/countries_combined_future_",estab_to_plot,"_v",vers,".pdf"),width = 7.5, height = 7)
      grid.draw(g)
      #dev.off()
      data_to_cor <- as.matrix(select(data_to_plot,x_to_plot,y_to_plot,size_to_plot))
      countries.impact.model <-
        impact.model <-
        lm(size_to_plot ~ x_to_plot + y_to_plot, data = data.frame(data_to_cor))
      countries.impact.cor <- cor.test(impact.model$model$size_to_plot,
                                       impact.model$fitted.values,
                                       method = method_cor)
    }
  }
} # end of the 2 loop for loop for plotting states and countries


```













