---
title: "Invasion Potential Alignment ALT"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{vignette-025-quadrant-plots-alt}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Build and Save Risk Correlation Quadrant Plots (Ensemble Extracts, Alternate Extraction Data)

# 1. Load required packages and set the same settings

```{r load required packages, warning=FALSE, message=FALSE}
library(slfrsk)
library(tidyverse)
library(ggrepel)
library(grid)
library(gridExtra)
```

```{r set to run different versions}
# Make transport scaled between one and zero
one_zero <- TRUE

# switch between present or future transport
present_transport <- FALSE

# Set what type of multiple correlation to calculate
method_cor <- "spearman"

# set whether to look at grape or wine production (PLOT LABELS ARE NOT CLEAN FOR WINE!!!)
#impact_prod <- "wine"
impact_prod <- "grapes"

# Plot the native countries?
native <- TRUE

# remove USA?
rem_usa <- TRUE

# NONFUNCTIONAL---Replace this with max mean quantile_0.50 quantile_0.75 quantile_0.90 to edit which metric to plot for establishment
#to run this, one needs to use different extract data sources further back in the cleaning process (vig01) by selecting a different variable
#estab_to_plot <- "obs_max" # max was used for the quadrant figures
estab_to_plot <- "obs_mean"
#estab_to_plot <- "quantile_0.50" #median
```

# 2. Load the full `suitability extract ensemble` data and the `summary_future` data (`states` and `countries`). 

Join the two dataframes and resolve conflicts. Use the `summary_future` data to lead the merging with `left_join()`.

  - mean suitability
  - median suitability
  
```{r read in data}
#suitability data load
data("states_extracts_ensemble")
data("countries_extracts_ensemble")

#summary data load
data("states_summary_future_ensemble")
data("countries_summary_future_ensemble")

#join the data
states_summary_future_ensemble <- states_summary_future_ensemble %>%
  left_join(., states_extracts_ensemble, by = "geopol_unit") %>%
  #drop grandmeanmax col, it is redundant
  dplyr::select(-grand_mean_max)

countries_summary_future_ensemble <- countries_summary_future_ensemble %>%
  left_join(., countries_extracts_ensemble, by = "geopol_unit") %>%
  #drop grandmeanmax col, it is redundant
  dplyr::select(-grand_mean_max)

#make it easier to port over parts of the vig-021 code by changing the names
states_summary_future <-  states_summary_future_ensemble
countries_summary_future <- countries_summary_future_ensemble
```

# 3. Drop the same `states` and `countries` that we dropped in the regular correlation plot.

```{r rm the same data as normal corplot}
#filter out some No Data for ensemble model
countries_summary_future <- countries_summary_future %>%
  filter(!ID %in% c("Antarctica", "Monaco", "Norfolk Island", "Spratly Islands"))

if(rem_usa){ # TRUE if you want to remove the USA
  rems <- c("Akrotiri and Dhekelia"
            ,"?land"
            ,"American Samoa"
            ,"Bouvet Island"
            ,"British Indian Ocean Territory"
            ,"Caspian Sea"
            ,"Christmas Island"
            ,"Clipperton Island"
            ,"Cocos Islands"
            ,"Falkland Islands"
            ,"Faroe Islands"
            ,"French Southern Territories"
            ,"Heard Island and McDonald Islands"
            ,"Mayotte"
            ,"Northern Mariana Islands"
            ,"Paracel Islands"
            ,"Pitcairn Islands"
            ,"Saint Pierre and Miquelon"
            ,"Tokelau"
            ,"United States Minor Outlying Islands"
            ,"Wallis and Futuna"
            ,"United States"
            ,"Philippines")
} else {
  rems <- c("Akrotiri and Dhekelia"
            ,"?land"
            ,"American Samoa"
            ,"Bouvet Island"
            ,"British Indian Ocean Territory"
            ,"Caspian Sea"
            ,"Christmas Island"
            ,"Clipperton Island"
            ,"Cocos Islands"
            ,"Falkland Islands"
            ,"Faroe Islands"
            ,"French Southern Territories"
            ,"Heard Island and McDonald Islands"
            ,"Mayotte"
            ,"Northern Mariana Islands"
            ,"Paracel Islands"
            ,"Pitcairn Islands"
            ,"Saint Pierre and Miquelon"
            ,"Tokelau"
            ,"United States Minor Outlying Islands"
            ,"Wallis and Futuna"
            ,"Philippines")

}

#do the rm now
countries_summary_future <- countries_summary_future %>% filter(!ID %in% rems)

#set the native countries
if(native){
  countries_summary_future$status[countries_summary_future$geopol_unit %in% c("China", "India", "Taiwan", "Vietnam")] <- "native"
} else {
  countries_summary_future$status[countries_summary_future$geopol_unit %in% c("China", "India", "Taiwan", "Vietnam")] <- "not established"
  }

```


# 4. Rescale the data the same way but also select the metric to look at.

```{r rescale states and isolate suit metric, do same for countries}
# scale the import data
if(one_zero){
  states_summary_future <- states_summary_future %>%
    # transport
    mutate(avg_infected_mass_scaled = log10(avg_infected_mass+1)) %>%
    mutate(avg_infected_mass_scaled = (avg_infected_mass_scaled - min(avg_infected_mass_scaled))) %>%
    mutate(avg_infected_mass_scaled = avg_infected_mass_scaled / max(avg_infected_mass_scaled)) %>%
    # establishment
    mutate(suitability_scaled = .[[estab_to_plot]])
} else {
  states_summary_future <- states_summary_future %>%
    # transport
    mutate(avg_infected_mass_scaled = log(avg_infected_mass+1)) %>%
    # establishment
    mutate(suitability_scaled = .[[estab_to_plot]])
}




if(one_zero){
  countries_summary_future <- countries_summary_future %>%
    # transport
    mutate(avg_infected_mass_scaled = log10(avg_infected_mass+1)) %>%
    mutate(avg_infected_mass_scaled = (avg_infected_mass_scaled - min(avg_infected_mass_scaled))) %>%
    mutate(avg_infected_mass_scaled = avg_infected_mass_scaled / max(avg_infected_mass_scaled)) %>%
    # establishment
    mutate(suitability_scaled = .[[estab_to_plot]])
} else {
  countries_summary_future <- countries_summary_future %>%
    # transport
    mutate(avg_infected_mass_scaled = log(avg_infected_mass+1)) %>%
    # establishment
    mutate(suitability_scaled = .[[estab_to_plot]])
}
```

# 5. Plotting settings

```{r params for plotting}
# label both of the countries (or states) with established SLF?
label.countries.est <- TRUE
label.states.est <- TRUE   #PRESENT: adding DE, MD, NJ  FUTURE: adding DE, MD, NJ, OH, WV
# labeling aesthetics for label.countries.est and label.states.est
nudge_countries.est.x <- c(-.05,0) 
nudge_countries.est.y <- c(.07,.1)
nudge_countries.est_future.x <- c(.05,-0.03)  #jpn and kor
nudge_countries.est_future.y <- c(.14,.05)
nudge_states.est_present.x <- c(0.00, -0.02,-0.04)  #PRESENT: adding DE, MD, NJ  
nudge_states.est_present.y <- c(0.10,-0.10,-0.10)
nudge_states.est_future.x <- c( 0.00,-0.10, 0.02,-0.04,0.00)  #FUTURE: adding OH, DE, MD, NJ, WV
nudge_states.est_future.y <- c(-0.10, 0.10,-0.17, 0.08,0.15)


# Choose quadrant intercepts
q_intercepts <- list(transport = 0.5, establishment = 0.5)
#set margins
hh = .001 
# Do not change this flips between plotting the states or plotting countries
plots_which <- c(TRUE,FALSE) 
# How many countries and states to highlight for labels?
top_to_plot <- 10 
#color aesthetics
risk_color <- "gray40"
risk_size <- 5

#constant looping var
i <- 1
```

# 6. Plotting code itself

```{r plot figures, warning=FALSE, message=FALSE, echo=TRUE, results="hide"}
for (i in 1:length(plots_which)) {
  
  #selects states or countries to plot
  statez = plots_which[i]
  
  ########################
  #DATA SELECTION STEP
  ########################

  #checks whether to use wine or grapes as the impact product to consider
  if(impact_prod == "wine"){
    if (statez) {
        data_to_plot <- states_summary_future
      
      #calls the exact variables to plot
      data_to_plot <- data_to_plot %>%
        mutate(
          x_to_plot = avg_infected_mass_scaled,
          y_to_plot = suitability_scaled,
          fill_to_plot = grapes,
          size_to_plot = avg_wine,
          color_to_plot = status
        ) %>%
        arrange(desc(grapes), (size_to_plot)) #so that the grape producers are on top
      
      #order the data that gets labeled based on top_to_plot
      data_to_label <- tail(data_to_plot, top_to_plot)  %>% arrange(desc(size_to_plot))
      #do welch t.test to assess if transport and establishment have the same relationship with grapes
      state_grape_t.test <- list(transport = t.test(avg_infected_mass_scaled~grapes, data = data_to_plot),
                                 establishment = t.test(suitability_scaled~grapes, data = data_to_plot))
      
      #add infected state labeling if turned on
      if(label.states.est){
        #FUTURE
          data_to_label_states.est <- data_to_plot %>%
          filter(geopol_unit %in% c("Delaware", "Maryland", "New Jersey", "Ohio", "West Virginia")) %>%
          arrange(desc(size_to_plot))
        data_to_label <- bind_rows(data_to_label,data_to_label_states.est)
      } #end of labeling for infected states
      
    } else {
      #selects the data to modify for plotting: COUNTRIES and selects the timing
        data_to_plot <- countries_summary_future
      
      data_to_plot <- data_to_plot %>%
        mutate(
          x_to_plot = avg_infected_mass_scaled,
          y_to_plot = suitability_scaled,
          fill_to_plot = grapes,
          size_to_plot = avg_wine,
          color_to_plot = status
        ) %>%
        arrange(desc(grapes), (size_to_plot)) #so that the grape producers are on top
      data_to_label <-
        tail(data_to_plot, top_to_plot)  %>%
        arrange(desc(size_to_plot)) %>%
        mutate(
          ID = recode(              #add ISO3 codes for the labeled countries (rather than all)
            ID,
            Italy = "ITA",
            France = "FRA",
            Spain = "ESP",
            China = "CHN",
            Argentina = "ARG",
            Chile = "CHL",
            Australia = "AUS",
            `South Africa` = "ZAF",
            Germany = "DEU",
            Portugal = "PRT"
          )
        )
      
      #do welch t.test to assess if transport and establishment have the same relationship with grapes
      country_grape_t.test <- list(transport = t.test(avg_infected_mass_scaled~grapes, data = data_to_plot),
                                   establishment = t.test(suitability_scaled~grapes, data = data_to_plot))
      if(label.countries.est){
        data_to_label_countries.est <- data_to_plot %>%
          filter(geopol_unit %in% c("Japan", "South Korea")) %>%
          mutate(ID = recode(ID,
                             Japan = "JPN",
                             `South Korea` = "KOR")) %>%
          arrange(desc(size_to_plot))
        data_to_label <- bind_rows(data_to_label,data_to_label_countries.est)
      }
    }
  } else if(impact_prod == "grapes"){
    if (statez) {
      #selects the data to modify for plotting: STATES and chooses the timing
        data_to_plot <- states_summary_future
      
      data_to_plot <- data_to_plot %>%
        mutate(
          x_to_plot = avg_infected_mass_scaled,
          y_to_plot = suitability_scaled,
          fill_to_plot = wine,
          size_to_plot = avg_prod, #avg_yield or avg_prod are the two options here for grapes
          color_to_plot = status
        ) %>%
        arrange(desc(grapes), (size_to_plot)) #so that the grape producers are on top
      #change the zeros to one's for plot size
      data_to_plot$size_to_plot[data_to_plot$size_to_plot == 0] <- 1

      data_to_label <-
        tail(data_to_plot, top_to_plot)  %>% arrange(desc(size_to_plot))
      state_grape_t.test <- list(transport = t.test(avg_infected_mass_scaled~grapes, data = data_to_plot),
                                 establishment = t.test(suitability_scaled~grapes, data = data_to_plot))
      
            #add infected state labeling if turned on
      if(label.states.est){
       #FUTURE
          data_to_label_states.est <- data_to_plot %>%
          filter(geopol_unit %in% c("Delaware", "Maryland", "New Jersey", "Ohio", "West Virginia")) %>%
          arrange(desc(size_to_plot))
        data_to_label <- bind_rows(data_to_label,data_to_label_states.est)
      } #end of labeling for infected states

    } else {
      #selects the data to modify for plotting: COUNTRIES and selects the timing
        data_to_plot <- countries_summary_future
      
      data_to_plot <- data_to_plot %>%
        mutate(
          x_to_plot = avg_infected_mass_scaled,
          y_to_plot = suitability_scaled,
          fill_to_plot = wine,
          size_to_plot = avg_prod, #avg_yield or avg_prod are the two options here for grapes
          color_to_plot = status
        ) %>%
        arrange(desc(grapes), (size_to_plot)) #so that the grape producers are on top
      #change the zeros to one's for plot size
      data_to_plot$size_to_plot[data_to_plot$size_to_plot == 0] <- 1

      data_to_label <-
        tail(data_to_plot, top_to_plot)  %>% arrange(desc(size_to_plot))
      
      data_to_label <- data_to_label %>%
        mutate(
          ID = recode(
            ID,
            Egypt = "EGY",
            Peru = "PER",
            India = "IND",
            Albania = "ALB",
            Iraq = "IRQ",
            Brazil = "BRA",
            Thailand = "THA",
            `South Africa` = "ZAF",
            China = "CHN",
            Armenia = "ARM",
            Italy = "ITA",
            Spain = "ESP",
            France = "FRA",
            Turkey = "TUR",
            Chile = "CHL",
            Argentina = "ARG",
            Australia = "AUS",
            `South Korea` = "KOR",
            Japan = "JPN"
          )
        )
      
      country_grape_t.test <- list(transport = t.test(avg_infected_mass_scaled~grapes, data = data_to_plot),
                                   establishment = t.test(suitability_scaled~grapes, data = data_to_plot))
      if(label.countries.est){
        data_to_label_countries.est <- data_to_plot %>%
          filter(geopol_unit %in% c("Japan", "South Korea")) %>%
          mutate(ID = recode(ID,
                             Japan = "JPN",
                             `South Korea` = "KOR")) %>%
          arrange(desc(size_to_plot))
        data_to_label <- bind_rows(data_to_label,data_to_label_countries.est)
      } # end of if label plotting JPN KOR
    }
  }


########################
#PLOT SETTINGS STEP
########################

#SET ALL OF THE NUDGES FOR LABELS
#need to add the conditional for present transport, since nudging only works based on the data
if(present_transport == FALSE){
  if(impact_prod == "wine"){
    #nudging
    if(statez) { #STATES FUTURE WINE
      nudge.x <- c(-.1,-.1,-.1,-.05,-.1,-0.05,-.1,.06,-.01,.05)
      nudge.y <- c(.2,.10,.10,.10,.07,.15,.15,-.07,-.12,-.12)    #CA, WA, NY, PA, OR, GA, OH, MI, VA, NC
      if(label.states.est){
        nudge.x <- c(nudge.x, nudge_states.est_future.x)
        nudge.y <- c(nudge.y, nudge_states.est_future.y)
      }
      ytitle <- "Establishment Potential"
    } else {  #COUNTRIES FUTURE WINE
      nudge.x <- c(-.3,-.27,-.05,-.26,.14,.1,.20,-.21,.08,.2)
      nudge.y <- c(.13, .095, .155, .09, -.11, .13, .15,.06,-.1,-.19)  #ITA, FRA, ESP, CHN, ARG, CHL, AUS, ZAF, DEU, PRT
      if(label.countries.est) {
        nudge.x <- c(nudge.x, nudge_countries.est_future.x) #JPN KOR
        nudge.y <- c(nudge.y, nudge_countries.est_future.y) #JPN KOR
      }
      ytitle <- "Establishment Potential"
    }
  } else if(impact_prod == "grapes"){
    #nudging
    if(statez) { #STATES FUTURE GRAPES
      nudge.x <- c(-.00,-.05,-.10,-.02, .10,-.05, .10, .00,-.05,-.025)   #yield: #CA, PA, WA, MI, NY, MO, OH, OR, GA, NC
      nudge.y <- c(-.10, .10, .08, .08,-.05, .09,-.05,-.10,-.15, .10)   #production: CA, WA, NY, PA, MI, OR, TX, VA, NC, MO
      if(label.states.est){
        nudge.x <- c(nudge.x, nudge_states.est_future.x)
        nudge.y <- c(nudge.y, nudge_states.est_future.y)
      }
      ytitle <- "Establishment Potential"
    } else {  #COUNTRIES FUTURE GRAPES
      #ensemble version
      # grape prod: CHN, ITA, ESP, FRA, TUR, IND, CHL, ARG, ZAF, AUS
      nudge.x <- c(-0.00, 0.01,-0.02, 0.00,-0.02, 0.03,-0.02, 0.08,-0.02, 0.06)
      nudge.y <- c(-0.10, 0.10, 0.15, 0.09,-0.24, 0.05, 0.13,-0.23, 0.29, -0.23)
      #old version
      # grape production: CHN, ITA, ESP, FRA, TUR, IND, CHL, ARG, ZAF, AUS
      #nudge.x <- c(-0.01,-0.00,-0.00,-0.04, 0.02,0.03,-0.04, 0.05,-0.03, 0.04)
      #nudge.y <- c(-0.10,-0.20,-0.27, 0.06,-0.24,0.07, 0.11,-0.23, 0.16,-0.10)
      if(label.countries.est) {
        nudge.x <- c(nudge.x, nudge_countries.est_future.x) #JPN KOR
        nudge.y <- c(nudge.y, nudge_countries.est_future.y) #JPN KOR
      }
      ytitle <- "Establishment Potential"
    }
  }
}
  

  #CREATE THE STATES PLOT
  (states_plot <- ggplot(data = data_to_plot) +
      #geom_text(data = risk_labels, mapping = aes(x = x, y = y,label = label), color = risk_color, size = risk_size) +
      geom_hline(yintercept = q_intercepts$establishment, linetype = "dashed") +
      geom_vline(xintercept = q_intercepts$transport, linetype = "dashed") +
      #  geom_abline(intercept = 0, slope = 1, linetype = "solid") + # add one to one
      geom_rect(mapping = aes(xmin=1.01, xmax=1.2, ymin=.49, ymax=.5), fill = "white") +
      geom_rect(mapping = aes(ymin=1.01, ymax=1.2, xmin=.49, xmax=.51), fill = "white") +
      geom_text_repel(data = data_to_label,
                      aes(x = x_to_plot, y = y_to_plot, label = ID),
                      min.segment.length = 0,
                      direction = "x",
                      nudge_y = nudge.y,
                      nudge_x = nudge.x
      ) +
      geom_point(
        aes(x = x_to_plot, y = y_to_plot, fill = fill_to_plot, size = size_to_plot, color = color_to_plot),
        shape = 21, stroke = 1.3, alpha = 0.75
      ) +
      scale_fill_manual(
        values = c("no" = "#ffffff", "yes" = "#C77CFF"),
        name = "Wine Production",
        labels = c("high", "low")
      ) +
      scale_color_manual(
        values = c(
          "red",
          "black",
          "blue"),
        breaks = c(
          "established",
          "not established",
          "native"),
        name = "Regional Status",
        labels = c("invaded","uninvaded", "native")
      ) +
      guides(
        fill = guide_legend(
          order = 1,
          override.aes = list(shape = 22, size = 5, alpha = 1)
        ),
        color = guide_legend(
          order = 3,
          override.aes = list(alpha = 1)
        ),
        size = guide_legend(    # Adjust size to edit legend grape production
          order = 2,            # circles.
          override.aes = list(size = c(1,3,6), alpha = 1)
        )
      ) +
      scale_size_continuous(name = "Grape Production",
                            trans = "log10",
                            range = c(1.5, 6),
                            breaks = c(1.5,2,6),
                            labels = c("low", "moderate", "high")
      ) +
      labs(x = "Transport Potential", y = ytitle) +
      ylim(0, 1.2) + xlim(0, 1.2) +
      theme(
        panel.grid.major = element_line(colour = "#f0f0f0"),
        panel.grid.minor = element_blank(),
        #panel.grid.major = element_blank(),
        #panel.border = element_rect(colour = "black", fill=NA, size=1),
        axis.line = element_line(colour = "black"),
        legend.position = c(0.7, 0.25),
        panel.background = element_blank(),
        plot.background = element_blank(),
        legend.background = element_blank(),#element_rect(colour = 'black', fill = 'white', linetype='solid'),
        axis.text = element_text(size = rel(1)),
        axis.title.x = element_text(hjust = .4),
        axis.title.y = element_text(hjust = .35),
        legend.title = element_text(face = "plain"),
        legend.key.size = unit(0.2, "cm"),
        legend.key = element_blank(),
        plot.margin = unit(c(hh, -5, hh, hh), units = "line"),
        axis.title = element_text(size = rel(1.3))
      ) +
      scale_x_continuous(breaks = c(0,.5,1),labels = c("low", "moderate", "high")) +
      scale_y_continuous(breaks = c(0,.5,1),labels = c("low", "moderate", "high")) + 
      coord_flip()
  )

  #get the legend
  get_legend <- function(myggplot) {
    tmp <- ggplot_gtable(ggplot_build(myggplot))
    leg <-
      which(sapply(tmp$grobs, function(x)
        x$name) == "guide-box")
    legend <- tmp$grobs[[leg]]
    return(legend)
  }
  #plot all of it
  if(statez) {
    legend_state <- legend <- get_legend(states_plot)
  } else {
    legend_country <- legend <- get_legend(states_plot)
  }


  (states_plot <- states_plot + theme(legend.position = "none"))


  #boxplot: ESTABLISHMENT
  (
    states_box_estab <- ggplot(data_to_plot) +
      geom_boxplot(
        aes(x = fill_to_plot, y = y_to_plot, group = fill_to_plot, fill = fill_to_plot),
        show.legend = FALSE,
        outlier.shape = NA,
        notch = FALSE,
        notchwidth = .25
      ) +
    ylim(0, 1) +
      scale_fill_manual(values = c("yes" = "#C77CFF", "no" = "#ffffff"), name = "") +
      theme(
        axis.line = element_blank(),
        axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks = element_blank(),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        legend.position = "none",
        panel.background = element_blank(),
        panel.border = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        plot.margin = unit(c(hh, hh, hh, hh), units = "line"),
        plot.background = element_blank()
      ) +
      coord_flip()
  )

  #boxplot: INTRODUCTION
  (
    states_box_intro <- ggplot(data_to_plot) +
      geom_boxplot(
        aes(x = fill_to_plot, y = x_to_plot, group = fill_to_plot, fill = fill_to_plot),
        show.legend = FALSE,
        outlier.shape = NA,
        notch = FALSE,
        notchwidth = .25
      ) +
      labs(x = "  ", y = NULL) +
      ylim(0, 1) +
      scale_fill_manual(
        values = c("yes" = "#C77CFF", "no" = "#ffffff"),
        name = ""
      ) +
      theme(
        axis.line = element_blank(),
        axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks = element_blank(),
        axis.title.x = element_blank(),
        legend.position = "none",
        panel.background = element_blank(),
        panel.border = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        plot.background = element_blank(),
        plot.margin = unit(c(hh, hh, hh, hh), units = "line")
      )
  )


  (
    states_plot_final <- states_plot +
      annotation_custom(
        grob = ggplotGrob(states_box_estab),
        xmin = 1.01,
        xmax = 1.11,
        ymin = -Inf,
        ymax = 1.05
      ) +
      # Insert ybp_grob inside the scatter plot
      annotation_custom(
        grob = ggplotGrob(states_box_intro),
        xmin = -.1,
        xmax = 1.05,
        ymin = 1,
        ymax = 1.1
      ) +

      annotation_custom(
        grob = rectGrob(gp = gpar(fill = "white", col = "white")),
        xmin = -1,
        xmax = 1.21,
        ymin = 1.1,
        ymax = 1.21
      ) +
      annotation_custom(
        grob = rectGrob(gp = gpar(fill = "white", col = "white")),
        ymin = -1,
        ymax = 1.21,
        xmin = 1.11,
        xmax = 1.21
      ) +
      scale_x_continuous(breaks = c(0,.5,1),labels = c("low", "moderate", "high"), limits = c(0,1.15)) +
      scale_y_continuous(breaks = c(0,.5,1),labels = c("low", "moderate", "high"), limits = c(0,1.15))
  )

  g <- ggplotGrob(states_plot_final)
  g$layout$clip[g$layout$name == "panel"] <- "off"
  #g$layout
  g$layout$z[g$layout$name == "panel"] <- 17 # Note that z for panel is 1.  Change it to something bigger.

#Correlation Analysis
  # present or future
 if(present_transport){
    if (statez) { #states present
      states_grob <- g
    } else {  #countries present
      countries_grob <- g
  } 
   }  else {    #future cases
    if (statez) { #states future
      states_grob <- g
    } else{ #countries future
      countries_grob <- g
      } 
            }

#SAVES RESULTS AS PDF
#plot the states+legend and save
if (statez) {
    #pdf(file.path(here::here(),"vignettes", paste0("states_combined_future_max_",impact_prod,".pdf")),width = 7.5, height = 7)
    grid.arrange(states_grob, legend_state, nrow=1, ncol = 2, widths = c(4,2), heights = 1)
    invisible(dev.off())
} else{
  #plot the countries+legend
    #pdf(file.path(here::here(),"vignettes", paste0("countries_combined_future_max_",impact_prod,".pdf")),width = 7.5, height = 7)
    grid.arrange(countries_grob, legend_country, nrow=1, ncol = 2, widths = c(4,2), heights = 1)
    invisible(dev.off())
}
  
} # end of the 2 loop for loop for plotting states and countries

```


# 7. Display the plots

```{r plot viz, fig.height= 12, fig.width=10, message=FALSE}
#set up a layout grid
  lay1 <- rbind(c( 1, 1, 1,NA,NA,NA),
               c( 1, 1, 1,NA,NA,NA),
               c( 1, 1, 1,NA, 3, 3),
               c(NA,NA,NA,NA, 3, 3),
               c( 2, 2, 2,NA, 3, 3),
               c( 2, 2, 2,NA,NA,NA),
               c( 2, 2, 2,NA,NA,NA)
  )

#nice version for knitting
grid.arrange(states_grob, countries_grob, legend_country, nrow = 7, ncol = 6, layout_matrix = lay1, heights = c(2,2,2, 0, 2,2,2), widths = c(1,1,1,0.1,0.6,0.6))
```

# 8. Get and print the correlations
```{r calculate correlations, warning=FALSE, message=FALSE}
#grapes
#STATES

#build model
states_grapes_impact_model <- lm(formula = log10_avg_prod ~ avg_infected_mass_scaled + suitability_scaled, data = states_summary_future)
#do cor
states_grapes_impact_cor <- cor.test(states_grapes_impact_model$model$log10_avg_prod, states_grapes_impact_model$fitted.values, method = method_cor)

#COUNTRIES
#build model
countries_grapes_impact_model <- lm(formula = log10_avg_prod ~ avg_infected_mass_scaled + suitability_scaled, data = countries_summary_future %>%
  filter(!ID %in% c("United States", 
                      "China", 
                      "India", 
                      "Taiwan", 
                      "Japan", 
                      "South Korea", 
                      "Vietnam")))
#do cor
countries_grapes_impact_cor <- cor.test(countries_grapes_impact_model$model$log10_avg_prod, countries_grapes_impact_model$fitted.values, method = method_cor)

#wine
#STATES
#model
states_wine_impact_model <- lm(formula = log10_avg_wine ~ avg_infected_mass_scaled + suitability_scaled, data = states_summary_future)
#cor
states_wine_impact_cor <- cor.test(states_wine_impact_model$model$log10_avg_wine, states_wine_impact_model$fitted.values, method = method_cor)

#COUNTRIES
#model
countries_wine_impact_model <- lm(formula = log10_avg_wine ~ avg_infected_mass_scaled + suitability_scaled, data = countries_summary_future %>%
  filter(!ID %in% c("United States", 
                      "China", 
                      "India", 
                      "Taiwan", 
                      "Japan", 
                      "South Korea", 
                      "Vietnam")))
#cor
countries_wine_impact_cor <- cor.test(countries_wine_impact_model$model$log10_avg_wine, countries_wine_impact_model$fitted.values, method = method_cor)
```

We can report the following alignment of potentials:

  1. **GRAPES**
    
  - U.S. State Alignment Correlation = `r formatC(states_grapes_impact_cor$estimate, format = "f", digits = 3)`, p = `r paste0(gsub('e', ' x 10^', formatC(states_grapes_impact_cor$p.value, format = "e", digits = 3)), '^')`
  - Country Alignment Correlation = `r formatC(countries_grapes_impact_cor$estimate, format = "f", digits = 3)`, p = `r paste0(gsub('e', ' x 10^', formatC(countries_grapes_impact_cor$p.value, format = "e", digits = 3)), '^')`
  
  2. **WINE**
    
  - U.S. State Alignment Correlation = `r formatC(states_wine_impact_cor$estimate, format = "f", digits = 3)`, p = `r paste0(gsub('e', ' x 10^', formatC(states_wine_impact_cor$p.value, format = "e", digits = 3)), '^')`
  - Country Alignment Correlation = `r formatC(countries_wine_impact_cor$estimate, format = "f", digits = 3)`, p = `r paste0(gsub('e', ' x 10^', formatC(countries_wine_impact_cor$p.value, format = "e", digits = 3)), '^')`
  
>Note: **`r impact_prod`** is the plotted impact product.
