---
title: "Trade and Regulatory Incident Relationship"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{vignette-042-trade-presence}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: inline
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Evaluate if Trade with Established States Relates to Regulatory Incidents or Establishment

***

This vignette fits a simple logistic regression model in the form of a binary dependent variable (either **states with regulatory incidents** or **states with established SLF**) on a continuous independent variable (trade with established states, mass in metric tons transformed as $log_{10}(value+1)$).

# Setup

## Load packages

```{r setup, warning=FALSE, results='hide', message=FALSE}
#packages to load
library(tidyverse) #data tidying
library(magrittr) #fast pipe assign
library(stargazer) #make easy model table for pub
library(sf) #manipulate by_county
library(here) #easy relative pathing

  #packages to use to extract raw data
#library(lubridate) #extract years
#library(tigris) #county level shapefiles
#library(lycormap) #package to get tinyslf dataset from
#options(tigris_use_cache = TRUE) # make sure that tigris does not keep downloading the files
```

## Read in trade data and established states

```{r read in data needed for functions}
#data to load

#slfrsk data (trade)
data("states_summary_present_ensemble", package = "slfrsk")

#recode the established variable as binary
states_summary_present_ensemble <- states_summary_present_ensemble %>%
  mutate(estab = ifelse(status == "established", 1, 0)) %>%
   dplyr::select(ID:status, estab, everything())
```

## Read in SLF reporting data

The data used here for the spread figure are based on an upcoming suite of packages (`lycodata` and `lycormap`) produced by the iEcoLab. The records in these packages contain data that are not yet available to the public, so all questions regarding the raw data should be directed to the creator of the `lycor*` suite, Seba De Bona (SDB, ([sebastiano.debona@temple.edu](mailto:sebastiano.debona@temple.edu))) or to Matthew R. Helmus ([mrhelmus@temple.edu](mailto:mrhelmus@temple.edu)). 

Notably these data are used only to get regulatory events. The `slfrsk` data already has established states, but it lacks the states that have regulatory events recorded but no established populations.

Here, we will read in a `.rda` file that was created using the example code below from the raw data. These example code chunks are adapted from SDB code.

```{r read in rda file for reg}
#data call
data(by_county, package = "slfrsk")
```

## Example code for raw data to county-level data

### Read in county shapefiles

We obtain the counties shapefile as a simple feature (sf) from the `tigris` package.

```{r counties shapefile, message=FALSE, eval=FALSE}
counties <- counties(cb = T, resolution = "5m")
```

Now, we read in the SLF records from the package `lycormap`. We also obtain the supplemental SLF records that contain the updated transportation data for uninvaded states, which was obtained from the raw data in the `lycodata` package. At the time of publication of `slfrsk`, additional data were saved as raw data that comes with `lycordata` (`slf_county_record_timeline.csv`).

```{r read in records data for map figure, eval=FALSE}
data("tinyslf", package = "lycormap")

# reading manually added data (for a later step)
man_county_data <- read_csv(file.path(here(), "..", "..", "lycordata", "data-raw", "additional", "slf_county_record_timeline.csv"),
                            col_types = cols(.default = "c", FIPS = "d", 
                                          RT_check = "l", ST_check = "l", RS_match = "l"))
```

### Merge and clean data

#### tinyslf

This code finds the county that each presence point corresponds to and attaches the county shapefile data to those rows. It also removes records that do not fall in a U.S. county (likely erroneous records). We also remove an errant record for Wasatch county, UT.

```{r tinyslf st combining and tidying, message=FALSE, eval=FALSE}
#adapted from SDB code in lycordata

#add column of temporary row ID's for later reference
tinyslf %<>% 
  add_column(row_ID = 1:nrow(.))

tinyslf %<>%
  # reducing data to coordinates only, and ID for future merging
  dplyr::select(latitude, longitude, row_ID) %>%
  # transforming into sf object
  st_as_sf(coords = c("longitude", "latitude"),
           crs = st_crs(counties)) %>% 
  #intersecting state polygons with data coordinates to match them to U.S. counties
  st_join(., counties, join = st_intersects) %>%
  #make it a tibble now
  as_tibble() %>% 
  #simplify to row ID, county name and identity
  dplyr::select(row_ID, county = NAME, GEOID) %>% 
  #then joining this into main tinyslf data
  left_join(tinyslf, ., by = "row_ID") %>% 
  #remove the temporary id column
  dplyr::select(-c(row_ID))

#remove the NA values for county. which are the points that do not fall in a county
tinyslf %<>%
  filter(!is.na(county))

#rm errant Wasatch co, UT record
tinyslf %<>%
  filter(!(state %in% "UT" & county %in% "Wasatch"))
```

Add the observations and establishment dates to the county outline data. Now, the earliest year of establishment or recording of an observation (dead or alive) are identified from the `tinyslf` dataset and then added accordingly to county shapefile data. Lastly, the projection is confirmed.

```{r find establishment and first records, message=FALSE, eval=FALSE}
#get the year of establishment
by_county <- tinyslf %>% 
  group_by(GEOID) %>% 
  arrange(year) %>% 
  filter(slf_established) %>% 
  summarize(YearOfEstablishment = min(year)) %>% 
  ungroup() %>%
  #join to the counties shapefile
  left_join(counties, ., by = "GEOID")

# and add the first record year
by_county <- tinyslf %>% 
  group_by(GEOID) %>% 
  arrange(year) %>% 
  filter(slf_present) %>% 
  summarize(FirstRecord = min(year)) %>% 
  ungroup() %>% 
  #join to shapefile that also includes year of establishment
  left_join(by_county, ., by = "GEOID")

# providing correct projections
by_county %<>% 
  st_transform('+proj=longlat +datum=WGS84')
```

#### slf_county_record_timeline.csv

Because our first record and establishment data are constantly updating from multiple sources, we have a googlesheet that is the source of `man_by_county` (saved as `slf_county_record_timeline.csv`). We again adapt code from SDB to tidy and splice these new records into the counties dataset.

```{r add manual counties data to counties data, message=FALSE, eval=FALSE}
#we can use the FIPS code to merge to the main data
#extract the status and the year of record/establishment
#there are three date columns that hopefully can be collapsed into 1

#standardize date format
man_county_data %<>% 
  mutate_at(vars(starts_with("Date")), .funs = ~parsedate::parse_date(.))

#data can have multiple dates associated to it, meaning a county can have info on when the first record was scored separately from establishment date
#not all infestations have an establishment data set, so cases where it is absent will be set to 2020 based on when data were analyzed
# for Date_Alive/Date_Morbound, we'll take the smallest, if both are present
man_county_data %<>% 
  mutate(Year_Alive = year(Date_Alive),
         Year_Morbund = year(Date_Morbund),
         Year_FirstRecord = ifelse(!(is.na(Year_Alive) & is.na(Year_Morbund)),
                                   pmin(Year_Alive, Year_Morbund, na.rm = T),
                                   NA)
         ) %>% 
  dplyr::select(-c(Year_Alive, Year_Morbund))

#create two variables that echo those in the by_county for years of first record and establishment
man_county_data %<>% 
  filter(!is.na(Status)) %>% 
  #first grabbing the date (if present) from the Date_Establish or Date_Alive/Morbound columns or setting 2020 to established records without dates
  #establishment
  mutate(YearOfEstablishment_man = ifelse(
    #if infestation and has a date of establishment
    Status == "Infestation" & !is.na(Date_Establish),
    year(Date_Establish),
    #if infestation and no date, sets to 2020
    ifelse(Status == "Infestation" & is.na(Date_Establish), 2020, NA)),
    #first record
    FirstRecord_man = ifelse(
    #not infested and has a year of first record
    Status != "Infestation" & !is.na(Year_FirstRecord),
    Year_FirstRecord,
    #infested and does not have a year of first record is set to 2020
    ifelse(Status == "Infestation" & is.na(Year_FirstRecord), 2020, NA))) %>% 
  dplyr::select(-Year_FirstRecord)
```

#### Merge the two datasets to just one county-level dataset

Now that both datasets are cleaned to have years of first record and establishment, they can be combined. A combined FIPS code is added to the full county dataset first.

```{r merge by_county and man_county_data, eval=FALSE}
#create combined FIPS columns
by_county %<>%
  mutate(FIPS = as.numeric(paste0(STATEFP, COUNTYFP)))

#merge
by_county <- man_county_data %>% 
  dplyr::select(FIPS, YearOfEstablishment_man, FirstRecord_man) %>% 
  left_join(by_county, ., by = "FIPS")
```

There may be some duplicate rows, and so the first record and establishment rows are retained. This dataset contains incomplete data for 2021, so we recode these as well, since our data are for as of 2020.
```{r pick first year rows, eval=FALSE}
#pick earliest year (between the records mined from the data and the manually compiled records)
by_county %<>% 
  mutate(YearOfEstablishment = pmin(YearOfEstablishment, YearOfEstablishment_man, na.rm = T),
         FirstRecord = pmin(FirstRecord, FirstRecord_man, na.rm = T)) %>%
  #recode the 2021 data
  #get cases where the first record is 2021, and set record to NA
  mutate(FirstRecord = case_when(
    FirstRecord == 2021 ~ NA_real_,
    TRUE ~ FirstRecord
  ),
  #get cases where the first record or year of establishment is 2021 and set establish to NA
  YearOfEstablishment = case_when(
    FirstRecord == 2021 ~ NA_real_,
    YearOfEstablishment == 2021 ~ NA_real_,
    TRUE ~ YearOfEstablishment
  )
  )
```

### Save the cleaned version of county-level data as `.rda`

`by_county` is now at a safe resolution for our use here. It is saved as `by_county.rda` for use in this plotting vignette.

```{r save by_county, eval=FALSE}
if(FALSE){
  save(by_county, file = file.path(here(), "data", "by_county.rda"))
}
```


# Get states with regulatory incidents

This current version uses the `data(by_country, package = "slfrsk")`. Again, `states_summary_present_ensemble` comes with established state data.
```{r new version with by_country}
#by_county version
reg_states <- by_county %>%
  st_drop_geometry() %>%
  as_tibble() %>%
  filter(!is.na(FirstRecord) | !is.na(YearOfEstablishment) | !is.na(FirstRecord_man) | !is.na(YearOfEstablishment_man)) %>%
  summarize(states = toupper(unique(STUSPS))) %>%
#clear it to just unique values
  distinct(states)

#recode the regulatory event states
states_summary_present_ensemble <- states_summary_present_ensemble %>%
  mutate(regulatory = ifelse(ID %in% reg_states$states, 1, 0)) %>%
  dplyr::select(ID:estab, regulatory, everything())

```

# Model the relationship

We have two models that we look at, specifically the relationship between trade with established states AND:

  1. establishment status as a binary (`estab`)
  2. regulatory event status (`regulatory`)
  
```{r fit the models}
#logistic regression
glm.fit1 <- glm(estab ~ log10_avg_infected_mass, data = states_summary_present_ensemble, family = "binomial")
#summarize to see fit
summary(glm.fit1)


#logistic regression for regulatory event states
glm.fit2 <- glm(regulatory ~ log10_avg_infected_mass, data = states_summary_present_ensemble, family = "binomial")
#summarize to see fit
summary(glm.fit2)

```

# Display Models

```{r stargazer output, results='asis'}
star_table <- stargazer(glm.fit1, glm.fit2, 
          type = "html", 
          intercept.bottom = T,
          ci = FALSE,
          digits = 2,
          style = "io",
          #style = "default",
          title = "SI Table: Logistic regression of SLF status on trade",
          out = paste0(here::here(),"/vignettes/logistic_models.doc")
          )
```

