---
title: "vignette-040-ee-data"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{vignette-040-ee-data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, results='hide', message=FALSE}
library(slfrsk)
library(tidyverse)
library(rgdal)
```

```{r read data}
data("states_summary_present")
data("states_summary_future")
data("countries_summary_present")
data("countries_summary_future")
```

We can pull country and state shapefiles from GADM
```{r read in the country and state shapefiles, message=FALSE, warning=FALSE}
  #shapefile of USA: data obtained from GADM: https://gadm.org/download_world.html
states <- readOGR(dsn = "/Volumes/GoogleDrive/Shared drives/slfRiskMapping/data/slfRisk/geo_shapefiles/gadm36_USA_shp/gadm36_USA_1.shp", verbose = T)

  #shapefile of world by countries
countries <- readOGR("/Volumes/GoogleDrive/Shared drives/slfRiskMapping/data/slfRisk/geo_shapefiles/gadm36_levels_shp/gadm36_0.shp", verbose = T, p4s = '+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0')

```

Clean the shape data and make future version copies of the shape data
```{r clean the shape data}
#STATES
#change DC name
#states@data$NAME_1 <- gsub(pattern = "Washington DC", replacement = "District of Columbia", x = states@data$NAME_1)

#COUNTRIES LAZY MAY WANT TO MAKE GSUB
countries@data$NAME_0[grep(pattern = "Ivoire", x = countries@data$NAME_0, value = F)] <- "Ivory Coast"
countries@data$NAME_0[grep(pattern = "ncipe", x = countries@data$NAME_0, value = F)] <- "Sao Tome and Principe"
countries@data$NAME_0[grep(pattern = "Cura", x = countries@data$NAME_0, value = F)] <- "Curacao"
countries@data$NAME_0[grep(pattern = "Saint Bart", x = countries@data$NAME_0, value = F)] <- "Saint Barthelemy"


#make a future version to hold the future data
states_future <- states
countries_future <- countries

```

Filter the shapefile data to be just the states and countries in the trade data
```{r filter the states to just those in the trade data}
#states
states <- states[states@data$NAME_1 %in% unique(states_summary_present$geopol_unit),]
states_future <- states_future[states_future@data$NAME_1 %in% unique(states_summary_future$geopol_unit),]
#countries
countries <- countries[countries@data$NAME_0 %in% unique(countries_summary_present$geopol_unit),]
countries_future <- countries_future[countries_future@data$NAME_0 %in% unique(countries_summary_future$geopol_unit),]

```

Now we can figure out how to join the trade data to the shapefiles (later to add the wine and grape production)
```{r add in the trade data}
#states
states@data <- left_join(states@data, states_summary_present, by = c("NAME_1" = "geopol_unit"))
states_future@data <- left_join(states_future@data, states_summary_future, by = c("NAME_1" = "geopol_unit"))

#countries
countries@data <- left_join(countries@data, countries_summary_present, by = c("NAME_0" = "geopol_unit"))
countries_future@data <- left_join(countries_future@data, countries_summary_future, by = c("NAME_0" = "geopol_unit"))

```

```{r rename fields}
#states
colnames(states@data)[13:31] <- c("gmean_max","gse_max","wine","grapes","avg_wine","se_wine","lavg_wine","avg_yield","se_yield","avg_prod","se_prod","lavg_yield","lavg_prod","avg_itrad","se_trad","avg_imass","se_mass","lavg_itrad","lavg_imass") 
colnames(states_future@data)[13:31] <- c("gmean_max","gse_max","wine","grapes","avg_wine","se_wine","lavg_wine","avg_yield","se_yield","avg_prod","se_prod","lavg_yield","lavg_prod","avg_itrad","se_trad","avg_imass","se_mass","lavg_itrad","lavg_imass") 

#countries
colnames(countries@data)[5:23] <- c("gmean_max","gse_max","wine","grapes","avg_wine","se_wine","lavg_wine","avg_yield","se_yield","avg_prod","se_prod","lavg_yield","lavg_prod","avg_itrad","se_trad","avg_imass","se_mass","lavg_itrad","lavg_imass") 
colnames(countries_future@data)[5:23] <- c("gmean_max","gse_max","wine","grapes","avg_wine","se_wine","lavg_wine","avg_yield","se_yield","avg_prod","se_prod","lavg_yield","lavg_prod","avg_itrad","se_trad","avg_imass","se_mass","lavg_itrad","lavg_imass") 

```

Write out the new shapefiles (may need to rename some cols in the `@data` field)
```{r write out the shapefiles, warning=FALSE, message=FALSE}
#states
writeOGR(obj = states, dsn = "/Volumes/GoogleDrive/Shared drives/slfRiskMapping/data/slfRisk/geo_shapefiles/gadm36_USA_with_data/",  driver="ESRI Shapefile", layer = "states", verbose = F)
writeOGR(obj = states_future, dsn = "/Volumes/GoogleDrive/Shared drives/slfRiskMapping/data/slfRisk/geo_shapefiles/gadm36_USA_with_data/",  driver="ESRI Shapefile", layer = "states_future", verbose = F)

#countries
writeOGR(obj = countries, dsn = "/Volumes/GoogleDrive/Shared drives/slfRiskMapping/data/slfRisk/geo_shapefiles/gadm36_levels_with_data/",  driver="ESRI Shapefile", layer = "countries", verbose = F)
writeOGR(obj = countries_future, dsn = "/Volumes/GoogleDrive/Shared drives/slfRiskMapping/data/slfRisk/geo_shapefiles/gadm36_levels_with_data/",  driver="ESRI Shapefile", layer = "countries_future", verbose = F)
```

```{r maybe test try plotting trade data}
#states_summary_present[match(x = states@data$NAME_1, table = states_summary_present$geopol_unit),]
#states@data

#states2 <- broom::tidy(states)

#fortifying to make points
states_gg <- fortify(states, region = "NAME_1")
states_gg <- left_join(states_gg, states_summary_present, by = c("id" = "geopol_unit"))
```

```{r try to test plot, cache=TRUE}
ggplot(states_gg[!states_gg$id %in% c("Alaska", "Hawaii"),]) +
  geom_polygon(aes(x = long, y = lat, group = group, fill = log10_avg_infected_mass)) +
  coord_quickmap()
```


