---
title: "SLF Spread Through Time"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{vignette-060-spread-figure}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: inline
bibliography: references.bib
link-citations: yes
---

```{r, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.align='left'
)
```

Map of SLF Spread Through Time in the US

***

We read in the packages to access data and generate the spread map. We also set the option to not download counties shapefiles every time (only the first).

# Setup

```{r setup, warning=FALSE, results='hide', message=FALSE}
library(slfrsk) #this package, has extract_enm()
library(lycormap) #package to get tinyslf dataset from
library(tidyverse)  #data manipulation
library(here) #making directory pathways easier on different instances
library(ggfortify) #fortify rasters
library(tigris) #county level shapefiles
library(sf) #simple features for easy shapefile manipulation
library(magrittr) #fast pipe assign
library(lubridate) #extract years
library(RColorBrewer) #palette
options(tigris_use_cache = TRUE) # make sure that tigris does not keep downloading the files
```

Now, we read in the SLF records from the package `lycormap`. We also obtain the supplemental SLF records that contain the updated transportation data for uninvaded states, which was obtained from the raw data in the `lycodata` package.

```{r read in records data for map figure}
data("tinyslf", package = "lycormap")

# reading manually added data (for a later step)
man_county_data <- read_csv(file.path("/Volumes/GoogleDrive-104607290079958926089/My Drive/spotted_lanternfly_ieco_projects/code/FINAL_CODE/R package/lycordata/", "data-raw", "additional",
                                      "slf_county_record_timeline.csv"),
                            col_types = cols(.default = "c", FIPS = "d", 
                                          RT_check = "l", ST_check = "l", RS_match = "l"))
```

We obtain the counties shapefile as a simple feature (sf) from the `tigris` package.

```{r counties shapefile, message=FALSE}
counties <- counties(cb = T, resolution = "5m")
```

# Merge and clean data

Here is the first of the code adapted from Seba De Bona (SDB), who built the `lycordata` and `lycormap` packages. Please email SDB if you are unable to access these packages ([sebastiano.debona@temple.edu](mailto:sebastiano.debona@temple.edu)). This code finds the county that  each presence point corresponds to and attaches the county shapefile data to those rows.

```{r tinyslf st combining and tidying, message=FALSE}
#adapted from SDB code in lycordata

#add column of temporary row ID's for later reference
tinyslf %<>% 
  add_column(row_ID = 1:nrow(.))

tinyslf %<>%
  # reducing data to coordinates only, and ID for future merging
  dplyr::select(latitude, longitude, row_ID) %>%
  # transforming into sf object
  st_as_sf(coords = c("longitude", "latitude"),
           crs = st_crs(counties)) %>% 
  #intersecting state polygons with data coordinates to match them to US counties
  st_join(., counties, join = st_intersects) %>%
  #make it a tibble now
  as_tibble() %>% 
  #simplify to row ID, county name and identity
  dplyr::select(row_ID, county = NAME, GEOID) %>% 
  #then joining this into main tinyslf data
  left_join(tinyslf, ., by = "row_ID") %>% 
  #remove the temporary id column
  dplyr::select(-c(row_ID))

#remove the NA values for county. which are the points that do not fall in a county
tinyslf %<>%
  filter(!is.na(county))

```

Add the observations and establishment dates to the county outline data
Now, the earliest year of establishment or recording of an observation (dead or alive) are identified from the `tinyslf` dataset and then added accordingly to county shapefile data. Lastly, the projection is confirmed.

```{r find establishment and first records, message=FALSE}
# defining the year of establishment
by_county <- tinyslf %>% 
  group_by(GEOID) %>% 
  arrange(year) %>% 
  filter(slf_established) %>% 
  summarize(YearOfEstablishment = min(year)) %>% 
  ungroup() %>% 
  left_join(counties, ., by = "GEOID")

# and adding the first record
by_county <- tinyslf %>% 
  group_by(GEOID) %>% 
  arrange(year) %>% 
  filter(slf_present) %>% 
  summarize(FirstRecord = min(year)) %>% 
  ungroup() %>% 
  left_join(by_county, ., by = "GEOID")

# providing correct projections
by_county %<>% 
  st_transform('+proj=longlat +datum=WGS84')

```

Because our first record and establishment data are constantly updating from multiple sources, we have a googlesheet that is the source of `man_by_county` (saved as `slf_county_record_timeline.csv`). We again adapt code from SDB to tidy and splice these new records into the counties dataset.

```{r add manual counties data to counties data, message=FALSE}

# we can use the FIPS code to merge to the main data
# now let's extract the status and the year of record/establishment
# there are three date columns that hopefully can be collapsed into 1
# first, let's check if all counties occur only once
stopifnot(
  man_county_data %>% 
    group_by(FIPS) %>% 
    tally() %>% 
    filter(n > 1) %>% 
    nrow(.) == 0
)

# now we check whether any row has more than one date entry

#standardize date format
man_county_data %<>% 
  mutate_at(vars(starts_with("Date")), .funs = ~parsedate::parse_date(.))

# now we check for overlaps in dates
man_county_data %>% 
  dplyr::select(starts_with("Date") & !ends_with("Descriptions")) %>% 
  mutate_all(~!is.na(.)) %>% 
  mutate(how_many_dates = apply(., MARGIN = 1, FUN = sum, na.rm = T)) %>% 
  filter(how_many_dates > 1) %>% 
  nrow(.) == 0

# according to the previous check, data can have multiple dates associated to it
# this means a county can have info on when the first record was scored separately from
# the establishment date. We need to keep this in mind

# let's now check whether all infestations have an establishment date associated to it
man_county_data %>% 
  filter(Status == "Infestation") %>% 
  head()

# It is immediately clear that the date is not always available.
# When not, we'll attribute to that datapoint 2020 as the latest 
# year establishement could have happened (given today is Jan 11 2021)
# for Date_Alive/Date_Morbound, we'll take the smallest, if both are present
man_county_data %<>% 
  mutate(Year_Alive = year(Date_Alive),
         Year_Morbund = year(Date_Morbund),
         Year_FirstRecord = ifelse(!(is.na(Year_Alive) & is.na(Year_Morbund)),
                                   pmin(Year_Alive, Year_Morbund, na.rm = T),
                                   NA)
         ) %>% 
  dplyr::select(-c(Year_Alive, Year_Morbund))

# we here create two variables that echo those in the by_county
man_county_data %<>% 
  filter(!is.na(Status)) %>% 
  # first grabbing the date (if present) from the Date_Establish or Date_Alive/Morbound columns
  # Then attributing 2020 to all other established records without date
  # doing this with nested ifelse statements
  mutate(YearOfEstablishment_man = ifelse(
    Status == "Infestation" & !is.na(Date_Establish),
    year(Date_Establish),
    ifelse(Status == "Infestation" & is.na(Date_Establish), 2020, NA)),
    FirstRecord_man = ifelse(
    Status != "Infestation" & !is.na(Year_FirstRecord),
    Year_FirstRecord,
    ifelse(Status == "Infestation" & is.na(Year_FirstRecord), 2020, NA))) %>% 
  dplyr::select(-Year_FirstRecord)

# for appropriate merging, we create a FIPS columns in the shapefiles
# then select only relevant columns from manual data
# then merge
by_county %<>%
  mutate(FIPS = as.numeric(paste0(STATEFP, COUNTYFP)))

# merging
by_county <- man_county_data %>% 
  dplyr::select(FIPS, YearOfEstablishment_man, FirstRecord_man) %>% 
  left_join(by_county, ., by = "FIPS")

# resolving conflict by picking the earliest year (between 
# the records mined from the data and the manually compiled records)
by_county %<>% 
  mutate(YearOfEstablishment = pmin(YearOfEstablishment, YearOfEstablishment_man, na.rm = T),
         FirstRecord = pmin(FirstRecord, FirstRecord_man, na.rm = T))

# transforming records and establishment
# (this will make coloring the leaflet map easier)
by_county %<>% 
  mutate(YearOfEstablishment = as.factor(YearOfEstablishment),
         FirstRecord = as.factor(FirstRecord))
```

# Plot spread figure
Lastly, we plot the data after identifying a set of focal counties to outline (this bypasses layering issues with `sf` objects and plotting with `ggplot2`).

```{r plot spread map, fig.dim=c(8,10)}
#get the focal counties to outline
focal_counties <- by_county %>%
  filter(!is.na(FirstRecord)) %>%
  as.data.frame(.) %>%
  dplyr::select(GEOID) %>%
  unlist()

map_plot <- ggplot() +
  #geom_sf(data = by_county %>% filter(!GEOID %in% focal_counties), fill = "#FFFFFF", size = 0.5) +
    geom_polygon(data = map_data('state'), aes(x = long, y = lat, group = group), fill = "#FFFFFF", color = "black", lwd = 0.10) +
  geom_polygon(data = map_data('state', c("Connecticut","Delaware", "Maryland", "New Jersey", "New York", "Ohio", "Pennsylvania", "Virginia", "West Virginia")), aes(x = long, y = lat, group = group), fill = NA, color = "#e31a1c", lwd = 0.75) +
  geom_sf(data = by_county %>% filter(GEOID %in% focal_counties), aes(color = FirstRecord, fill = YearOfEstablishment), show.legend = T) +
  coord_sf(xlim = c(-124.7628, -66.94889), ylim = c(24.52042, 49.3833), expand = FALSE) +
  labs(x = "", y = "") +
  theme_bw() +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_rect(fill = "aliceblue")) +
  scale_color_brewer(palette = "YlOrRd", direction = 1, name = "Regulatory Incident", guide = guide_legend(override.aes = list(fill = "#808080"))) +
  scale_fill_brewer(palette = "YlOrRd", direction = 1, name = "Establishment", na.value = "#808080", breaks = c(2014,2015,2016,2017,2018,2019,2020)) +
  ggtitle(label = "")

map_plot
```
# Save plot output

```{r save spread plot for pub, eval=FALSE}
pdf(file.path(here::here(),"vignettes", paste0("spread_figure_v3.pdf")), width = (18/2.54), height = (14/2.54))
map_plot
invisible(dev.off())

```

# Extras

We also include a modification of SDB code here to show how to make a nice `leaflet` interactive map too... but we do not use this code.
```{r optional leaflet plotting, eval=TRUE, fig.dim=c(8,10)}
library(leaflet)

colpal <- colorFactor(c(brewer.pal(7, "YlOrRd")),
                   by_county$YearOfEstablishment,
                   na.color = "transparent")
fillpal <- colorFactor(c(brewer.pal(7, "YlOrRd")),
                   by_county$FirstRecord,
                   na.color = "transparent")

leaflet(data = by_county, 
        options = leafletOptions(minZoom = 4, maxZoom = 18)) %>% 
  addProviderTiles(providers$CartoDB.Positron) %>% 
  addPolygons(label = ~NAME,
              color = ~colpal(FirstRecord),
              weight = 2,
              smoothFactor = 0.5,
              opacity = 0.5,
              fillOpacity = 0.5,
              fillColor = ~fillpal(YearOfEstablishment)) %>% 
  setView(lat = 40.80, lng = -100, zoom = 4) %>% 
  addLegend("bottomleft", pal = fillpal,
    title = "Invaded since",
    values = ~YearOfEstablishment,
    opacity = 1
  )

```



