---
title: "Risk Map Plotting"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{vignette-030-risk-maps}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Build and Save Maps of Transport and Establishment Potential

***

This vignette produces the published map figures that demonstrate where **SLF** is most likely to become established, the states and countries that trade most with the `future` scenario infected states, and shows locations of important viticultural areas (areas of high **impact** potential).

# Setup

Here we load the packages required to visualize the maps.

```{r setup, warning=FALSE, results='hide', message=FALSE}
library(slfrsk)
library(tidyverse)
library(rgeos)
library(scales)
```

## Toggle plotting options

As in `vignette-021-quadrant-plots`, we have several parameters that can be modified to visualize different versions of the figures. Here is where one could change the version of data for the map figures via the following: 

  1. present_transport: Use `present` or `future` infected U.S. states? 
      - default is `FALSE` for `future`
  2. top_to_plot: Set the number of top **transport** potential partners to plot arrows for?
      - default is `10`

```{r plotting options}
#we need a switch for this map that does the same thing for including the following states: New York, Ohio, West Virginia, North Carolina
present_transport <- FALSE
# How many countries and states to highlight for labels?
top_to_plot <- 10 
```

To build the map plots, we have several sources of data that we need to read in, namely: 

  1. viticultural area geographic locations
  2. suitability maps for `U.S. states` and all `countries` from the ensemble model
  3. locations to point to trade partners (capital cities of `states` and `countries`)
  
Of note, the actual trade data (**transport** potential) will be read conditionally in a later chunk below.

```{r read in the data we need}
#location of wineries (US and COUNTRIES)
data("wineries")
#need usa enviro raster
data("suitability_usa_df")
data("suitability_countries_df")
#state centers for plot
data("states_centers")
data("countries_centers")

```

# Data preparation

Now that we have most of the data loaded, we can go ahead and conditionally load the **transport** potential data based on the state of `present_transport`. The **transport** potential data for `states` and `countries` are then combined with the appropriate trade partner coordinate data and an approximate center of the **SLF** invasion in the U.S. is added as an origin for depicting trade arrows in the map figures. The `states` data are presented first and then the `countries` thereafter.

## States

```{r states trade, message=FALSE}
#need states trade data
#need to read in trade data to shape network in map figure
if(present_transport == TRUE){
  data("states_summary_present")
  states_centers <- states_centers %>%
    #remove the invaded states
    filter(!geopol_unit %in% c("Delaware", "Maryland", "New Jersey", "Pennsylvania", "Virginia")) %>%
    #remove the others in the dataset
    filter(!is.na(x), !is.na(y)) %>%
    rbind(c("SLF", NA, 40.400189, -75.917622)) %>%
    mutate(x = as.numeric(x), y = as.numeric(y))
  
#now add the "invaded core" == SLF as a column
states_summary_present <- states_summary_present %>%
  cbind("SLF", .) %>%
  as_tibble() %>%
  mutate(origin = `"SLF"`) %>%
  dplyr::select(-`"SLF"`) %>%
  dplyr::select(origin, everything())

#now merge the coords for destinations (ONLY WINE PRODUCING STATES)
states_trade_final <- left_join(states_summary_present, states_centers, by = "geopol_unit") %>%
  filter(!is.na(x), !is.na(y))

#add the coords for SLF as a separate set of cols
states_trade_final <- states_centers %>%
  filter(geopol_unit == "SLF") %>%
  dplyr::select(x,y) %>%
  mutate(xend = x, yend = y) %>%
  dplyr::select(xend, yend) %>%
  cbind(states_trade_final, .) %>%
  as_tibble()
  
} else {
  data("states_summary_future")
  
    states_centers <- states_centers %>%
    #remove the invaded states
    filter(!geopol_unit %in% c("Delaware", "Maryland", "New Jersey", "New York", "North Carolina", "Ohio", "Pennsylvania", "Virginia", "West Virginia")) %>%
    #remove the others in the dataset
    filter(!is.na(x), !is.na(y)) %>%
    rbind(c("SLF", NA, 40.400189, -75.917622)) %>%
    mutate(x = as.numeric(x), y = as.numeric(y))

        #now add the "invaded core" == SLF as a column
    states_summary_future <- states_summary_future %>%
      cbind("SLF", .) %>%
      as_tibble() %>%
      mutate(origin = `"SLF"`) %>%
      dplyr::select(-`"SLF"`) %>%
      dplyr::select(origin, everything())
    
    #now merge the coords for destinations (ONLY WINE PRODUCING STATES)
    states_trade_final <- left_join(states_summary_future, states_centers, by = "geopol_unit") %>%
      filter(!is.na(x), !is.na(y))
    
    #add the coords for SLF as a separate set of cols
    states_trade_final <- states_centers %>%
      filter(geopol_unit == "SLF") %>%
      dplyr::select(x,y) %>%
      mutate(xend = x, yend = y) %>%
      dplyr::select(xend, yend) %>%
      cbind(states_trade_final, .) %>%
      as_tibble()

    }

```

## Countries

```{r countries trade, message=FALSE}

#edit the world centers to also have the SLF US invasion point
countries_centers <- countries_centers %>%
  #rm the others in the dataset
  filter(!is.na(x), !is.na(y)) %>%
  #add center for SLF invasion in US
  #rbind(c("SLF", NA, -75.917622, 40.400189)) %>%
  rbind(c(rep("SLF", times = 4), -75.917622, 40.400189)) %>%  #(x,y) for coords
  #add the netherlands, not a wine country but needs a set of coords
  #rbind(c("Netherlands", NA, unlist(gCentroid(SpatialPointsDataFrame(map_data('world')[map_data('world')$region %in% "Netherlands",1:2], data = map_data('world')[map_data('world')$region %in% "Netherlands",]))@coords))) %>%
  #same for indonesia
  #rbind(c("Indonesia", NA, unlist(gCentroid(SpatialPointsDataFrame(map_data('world')[map_data('world')$region %in% "Indonesia",1:2], data = map_data('world')[map_data('world')$region %in% "Indonesia",]))@coords))) %>%
  #same for these if going for top 25: Singapore, Taiwan, Saudi Arabia, Vietnam, Venezuela, Gibraltar
  #mutate(x = as.numeric(x), y = as.numeric(y), avg_wine = as.numeric(avg_wine))
  mutate(x = as.numeric(x), y = as.numeric(y))

#now we manually bump Canada's coords up a little bit to make more visible
countries_centers$y[countries_centers$geopol_unit == "Canada"] <- countries_centers$y[countries_centers$geopol_unit == "Canada"] + 5


if(present_transport == TRUE){
  data("countries_summary_present")
  
  #filter out the countries with SLF
  countries_summary_present <- countries_summary_present %>%
  filter(!geopol_unit %in% c("China", "Japan", "Korea, South", "India")) %>%
  cbind("SLF", .) %>%
  as_tibble() %>%
  mutate(origin = `"SLF"`) %>%
  dplyr::select(-`"SLF"`) %>%
  dplyr::select(origin, everything())
  
  #now merge the coords for destinations (ONLY WINE PRODUCING countries)
countries_trade_final <- left_join(countries_summary_present, countries_centers, by = "geopol_unit") %>%
  filter(!is.na(x), !is.na(y))


} else {
  data("countries_summary_future")
  
  #filter out the countries with SLF
  countries_summary_future <- countries_summary_future %>%
  filter(!geopol_unit %in% c("China", "Japan", "Korea, South", "India")) %>%
  cbind("SLF", .) %>%
  as_tibble() %>%
  mutate(origin = `"SLF"`) %>%
  dplyr::select(-`"SLF"`) %>%
  dplyr::select(origin, everything())

    #now merge the coords for destinations (ONLY WINE PRODUCING countries)
countries_trade_final <- left_join(countries_summary_future, countries_centers) %>%
  filter(!is.na(x), !is.na(y))
  
}

#add the coords for SLF as a separate set of cols
countries_trade_final <- countries_centers %>%
  filter(geopol_unit == "SLF") %>%
  dplyr::select(x,y) %>%
  mutate(xend = x, yend = y) %>%
  dplyr::select(xend, yend) %>%
  cbind(countries_trade_final, .) %>%
  as_tibble()

```

#  Plots

We are now able to plot both the `states` and `countries` published map figures. We stepwise add in the different features, beginning with the suitability raster and geopolitical boundary outlines (the latter ships with `ggplot2` in `map_data()`). We then add <span style="color: red;">red outlines</span> for the U.S. states that have established **SLF** according to the scenario selected with `present_transport`. Following that, <span style="color:#a020f0;">purple points</span> are added for viticultural areas. Lastly, we add arrows from the approximate center of the **SLF** invasion in the U.S. to high **transport** potential trade partners.

## States

```{r the US map, fig.height= 5, fig.width=8}

#plot just the states and suitability first
map_plot <- ggplot() +
  geom_raster(data = suitability_usa_df, aes(x = x, y = y, fill = rescale(value)), alpha=0.9, show.legend = T) +
  scale_fill_gradientn(limits= c(0,1), name = "Suitability", colors = rev(c("#e31a1c","#fd8d3c", "#fecc5c", "#FFFFFF"))) +
  geom_polygon(data = map_data('state'), aes(x = long, y = lat, group = group), fill = NA, color = "black", lwd = 0.10) +
  theme_bw() +
  labs(x = "", y = "") +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank()) +
  coord_quickmap(xlim = c(-124.7628, -66.94889), ylim = c(24.52042, 49.3833)) +
  ggtitle(label = "")

#add in lines for the infected states
if(present_transport == TRUE){
  map_plot <- map_plot +
    geom_polygon(data = map_data('state', c("Delaware", "Maryland", "New Jersey", "Pennsylvania", "Virginia")), aes(x = long, y = lat, group = group), fill = NA, color = "#e31a1c", lwd = 0.75)
  
} else{
  map_plot <- map_plot +
    geom_polygon(data = map_data('state', c("Delaware", "Maryland", "New Jersey", "New York", "North Carolina", "Ohio", "Pennsylvania", "Virginia", "West Virginia")), aes(x = long, y = lat, group = group), fill = NA, color = "#e31a1c", lwd = 0.75)
}

#add in the wine AVA data
map_plot <- map_plot +
    geom_point(data = wineries %>% filter(Country == "United States"), aes(x = x, y = y), color = "black", fill = "purple", size = 1.5, shape = 21, alpha = 0.99, show.legend = F)

#add in the trade
map_plot <- map_plot +
    geom_curve(aes(x = x, y = y, xend = xend, yend = yend), data = states_trade_final %>%  arrange(desc(avg_infected_mass)) %>% .[1:top_to_plot,], size = 1, curvature = 0.33, alpha = 0.8, show.legend = T, color = "#778899",
               arrow = arrow(ends = "first", length = unit(0.01, "npc"), type = "closed"))

#show it
map_plot

```

## Countries

```{r plot map of countries, warning=FALSE, fig.height= 9*.75, fig.width=16*.75}
#plot just the countries and suitability first
map_plot_countries <- ggplot() +
  geom_raster(data = suitability_countries_df, aes(x = x, y = y, fill = rescale(value)), alpha=0.9, show.legend = T) +
  scale_fill_gradientn(limits= c(0,1), name = "Suitability", colors = rev(c("#e31a1c","#fd8d3c", "#fecc5c", "#FFFFFF"))) +
  geom_polygon(data = map_data('world'), aes(x = long, y = lat, group = group), fill = NA, color = "black", lwd = 0.15) +
  theme_bw() +
  labs(x = "", y = "") +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank()) +
  coord_quickmap(xlim = c(-164.5, 163.5), ylim = c(-55,85)) +
  ggtitle(label = "")
  
#add in lines for the infected states
if(present_transport == TRUE){
  map_plot_countries <- map_plot_countries +
    geom_polygon(data = map_data('state', c("Delaware", "Maryland", "New Jersey", "Pennsylvania", "Virginia")), aes(x = long, y = lat, group = group), fill = NA, color = "red", lwd = 0.5)
} else{
  map_plot_countries <- map_plot_countries +
    geom_polygon(data = map_data('state', c("Delaware", "Maryland", "New Jersey", "New York", "North Carolina", "Ohio", "Pennsylvania", "Virginia", "West Virginia")), aes(x = long, y = lat, group = group), fill = NA, color = "#e31a1c", lwd = 0.75)
}
  

#add in the wine AVA data (global equivalent)
map_plot_countries <- map_plot_countries +
  geom_point(data = wineries, aes(x = x, y = y), color = "black", fill = "purple", size = 0.5, shape = 21, alpha = 0.99, stroke = 0.2, show.legend = F)

#add in the trade
map_plot_countries <- map_plot_countries +
  geom_curve(aes(x = x, y = y, xend = xend, yend = yend), data = countries_trade_final %>%  arrange(desc(avg_infected_mass)) %>% .[1:top_to_plot,], size = 0.45, curvature = 0.33, alpha = 0.8, show.legend = T, color = "#778899", arrow = arrow(ends = "first", length = unit(0.01, "npc"), type = "closed")) 

#print it
map_plot_countries 
```

# Save plot output

Lastly, we can save the output of our map figure plots as `.pdf` files to `/vignettes`. Note that this chunk does not run by default but can be run by setting the top `if` from `FALSE` to `TRUE`!

```{r save map plots, eval=FALSE, warning=FALSE, message=FALSE}
if(FALSE){
  if(present_transport){
#states
  ggsave(plot = map_plot, filename = file.path(here::here(),"vignettes", paste0("states_map_present.pdf")), width = 7.5, height = 6.5, dpi = 300)
#countries
  ggsave(plot = map_plot_countries, filename = file.path(here::here(),"vignettes", paste0("countries_map_present.pdf")), width = 7.5, height = 6.5, dpi = 300)
} else{
  #states
  ggsave(plot = map_plot, filename = file.path(here::here(),"vignettes", paste0("states_map_future.pdf")), width = 7.5, height = 6.5, dpi = 300)
#countries
  ggsave(plot = map_plot_countries, filename = file.path(here::here(),"vignettes", paste0("countries_map_future.pdf")), width = 7.5, height = 6.5, dpi = 300)
}

}

```
