---
title: "Trade and Regulatory Incident Relationship"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{vignette-042-trade-presence}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: inline
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Evaluate if Trade with Established States Relates to Regulatory Incidents or Establishment

***

This vignette fits a simple logistic regression model in the form of a binary dependent variable (either **states with regulatory incidents** or **states with established SLF**) on a continuous independent variable (trade with established states, mass in metric tons transformed as $log_{10}(value+1)$).

# Setup

## Load packages

```{r setup, warning=FALSE, results='hide', message=FALSE}
#packages to load
library(tidyverse) #data tidying
library(magrittr) #fast pipe assign
library(lubridate) #extract years
library(stargazer) #make easy model table for pub
library(sf) #manipulate by_county
library(here) #easy relative pathing
```

## Read in data

```{r read in data needed for functions}
#data to load
#by county reg and estab
data(by_county, package = "slfrsk")

#slfrsk data (trade)
data("states_summary_present_ensemble", package = "slfrsk")

#recode the established variable as binary
states_summary_present_ensemble <- states_summary_present_ensemble %>%
  mutate(estab = ifelse(status == "established", 1, 0)) %>%
   dplyr::select(ID:status, estab, everything())
```

```{r}
#lycormap and lycordata data
#separate version with regulatory incidents
data("tinyslf", package = "lycormap")

# reading manually added data (for a later step)
man_county_data <- read_csv(file.path(here(), "..", "..", "lycordata", "data-raw", "additional", "slf_county_record_timeline.csv"),
                            col_types = cols(.default = "c", FIPS = "d", 
                                             RT_check = "l", ST_check = "l", RS_match = "l"))
```

## Quick clean up of `lycordata`

We now use some other data to get states with regulatory events. These data come from the `lycordata` package. Please email SDB if unable to access the packages/data ([sebastiano.debona@temple.edu](mailto:sebastiano.debona@temple.edu)). The `slfrsk` data already has established states, but it lacks the states that have regulatory events recorded but no established populations.

```{r tidying and summarizing the regulatory states}
#standardize the dates in the data
man_county_data %<>% 
  mutate_at(vars(starts_with("Date")), .funs = ~parsedate::parse_date(.))

#get years for the manual data
man_county_data %<>% 
  mutate(Year_Alive = year(Date_Alive),
         Year_Morbund = year(Date_Morbund),
         Year_FirstRecord = ifelse(!(is.na(Year_Alive) & is.na(Year_Morbund)),
                                   pmin(Year_Alive, Year_Morbund, na.rm = T),
                                   NA)
  ) %>% 
  dplyr::select(-c(Year_Alive, Year_Morbund))

# we here create two variables that echo those in the by_county
man_county_data %<>% 
  filter(!is.na(Status)) %>% 
  # first grabbing the date (if present) from the Date_Establish or Date_Alive/Morbound columns
  # Then attributing 2020 to all other established records without date
  # doing this with nested ifelse statements
  mutate(YearOfEstablishment_man = ifelse(
    Status == "Infestation" & !is.na(Date_Establish),
    year(Date_Establish),
    ifelse(Status == "Infestation" & is.na(Date_Establish), 2020, NA)),
    FirstRecord_man = ifelse(
      Status != "Infestation" & !is.na(Year_FirstRecord),
      Year_FirstRecord,
      ifelse(Status == "Infestation" & is.na(Year_FirstRecord), 2020, NA))) %>% 
  dplyr::select(-Year_FirstRecord)

#get the states by pulling counties with records or establishment
reg_states <- man_county_data %>%
  rename(state = State) %>%
  filter(!is.na(FirstRecord_man) | !is.na(YearOfEstablishment_man)) %>%
  summarize(states = toupper(unique(state))) %>%
#then bind to tinyslf main data to it
  bind_rows(tinyslf %>%
              filter(slf_present == TRUE | slf_established == TRUE) %>%
              summarize(states = unique(state)), .) %>%
#clear it to just unique values
  distinct(states)

#by_county version
reg_states <- by_county %>%
  st_drop_geometry() %>%
  as_tibble() %>%
  filter(!is.na(FirstRecord) | !is.na(YearOfEstablishment) | !is.na(FirstRecord_man) | !is.na(YearOfEstablishment_man)) %>%
  summarize(states = toupper(unique(STUSPS))) %>%
#clear it to just unique values
  distinct(states)

#recode the regulatory event states
states_summary_present_ensemble <- states_summary_present_ensemble %>%
  mutate(regulatory = ifelse(ID %in% reg_states$states, 1, 0)) %>%
  dplyr::select(ID:estab, regulatory, everything())

```

# Model the relationship

We have two models that we look at, specifically the relationship between trade with established states AND:

  1. establishment status as a binary (`estab`)
  2. regulatory event status (`regulatory`)
  
```{r fit the models}
#logistic regression
glm.fit1 <- glm(estab ~ log10_avg_infected_mass, data = states_summary_present_ensemble, family = "binomial")
#summarize to see fit
summary(glm.fit1)


#logistic regression for regulatory event states
glm.fit2 <- glm(regulatory ~ log10_avg_infected_mass, data = states_summary_present_ensemble, family = "binomial")
#summarize to see fit
summary(glm.fit2)

```

# Display Models

```{r stargazer output, results='asis'}
star_table <- stargazer(glm.fit1, glm.fit2, 
          type = "html", 
          intercept.bottom = T,
          ci = FALSE,
          digits = 2,
          style = "io",
          #style = "default",
          title = "SI Table: Logistic regression of SLF status on trade",
          out = paste0(here::here(),"/vignettes/logistic_models.doc")
          )
```

